### 2019-09-11 modeling_poc.Rmd
Adding report markdown and supporting modeling file.

---
title: 'Beta to Release Matching: Modeling'
author: "Corey Dow-Hygelund"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## tl;dr

[Statistical matching](https://en.wikipedia.org/wiki/Matching_(statistics)) methods were employed to investigate the posssibility of finding a subset of Beta clients that are representative of Release. The specific use-case of this proof-of-concept is utilizing engagement, configuration, and environment covariates of the clients for matching. Validation of the matching is performed on a hold-out set of Firefox performance covariates. Such a a use-case has real-world analogies.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
source('supporting_funcs.R')
```

## Data Loading/Preparation
```{r}
# define response field
label <- 'is_release'

# load up dataset
df <- read.csv('data/4week/US/df.csv') %>%
  select(-c(CONTENT_PAINT_TIME_PARENT, COMPOSITE_TIME_CONTENT, COMPOSITE_TIME_PARENT, CONTENT_FRAME_TIME_PARENT)) %>% # All NAs 
  select(-e10s_enabled) %>% # All True
  mutate(default_search_engine = as.factor(normalize_search_engine(default_search_engine))) %>%
  mutate(profile_age_cat = as.factor(normalize_profile_age(profile_age))) %>%
  mutate(distro_id_norm = as.factor(normalize_distro_id(distribution_id))) %>%
  mutate(timezone_cat = as.factor(normalize_timezone(timezone_offset))) %>%
  mutate(memory_cat = as.factor(normalize_memory(memory_mb))) %>% 
  mutate(cpu_speed_cat = as.factor(normalize_cpu_speed(cpu_speed_mhz))) %>% 
  mutate(cpu_cores_cat = as.factor(normalize_cpu_cores(cpu_cores))) %>% 
  mutate(!!label := case_when(
    label == 'beta' ~ FALSE,
    TRUE ~ TRUE)) %>%
  filter(startup_ms >= 0 & profile_age >= 0)
  # filter(country %in% c('US', 'GB'))

# complete cases (specifically targets perf metrics)
df_c <- df %>% na.omit()
```

```{r}
# Various sampling for beta overrepresentation
df_beta <- df_c %>% filter(is_release == FALSE)
df_rel <- df_c %>% filter(is_release == TRUE)
n_beta <- nrow(df_beta)

build_df <- function(multiple){
  df <- df_rel %>%
    sample_n(size = round(n_beta / multiple)) %>%
    rbind(df_beta)
}

df_1x <- build_df(1)
df_2x <- build_df(2)
df_4x <- build_df(4)

```

Downsample to allow nearest neighbors to complete within the year. 
```{r}
df_1x_sm <- df_1x %>% sample_n(size = 70000)
df_2x_sm <- df_2x %>% sample_n(size = 70000)
df_4x_sm <- df_4x %>% sample_n(size = 70000)
```

## Covariates

Define the covariates for modeling and validation.

```{r}
cont_log_covariates <- c('active_hours', 'daily_max_tabs', 'daily_tabs_opened', 'search_count', 'daily_unique_domains',
                     'num_bookmarks', 'num_addons', 'num_active_days', 'num_pages', 'uri_count', 'session_length')

cont_covariates <- c('profile_age', 'cpu_cores', 'cpu_speed_mhz', 'daily_num_sessions_started', 'timezone_offset')

cat_covariates <- c('country', 'timezone_cat', 'distribution_id', 'profile_age_cat')

holdout_covariates <- c("CONTENT_PAINT_TIME_CONTENT", "TIME_TO_DOM_CONTENT_LOADED_END_MS", 
                        "MEMORY_TOTAL", "TIME_TO_DOM_COMPLETE_MS", "FX_PAGE_LOAD_MS_2_PARENT", 
                        "FX_TAB_SWITCH_TOTAL_E10S_MS",
                        "CONTENT_FRAME_TIME_GPU", "COMPOSITE_TIME_GPU",
                        "TIME_TO_LOAD_EVENT_END_MS", 'startup_ms', 'content_crashes')
```

## Variable Selection
None performed at the moment. Literature suggests that using those related to outcome variables.

## Experiments
Groups of models to train. Complete alchemy at this point...

```{r}
exp_1 <- c('uri_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat', 'is_wow64', 'fxa_configured', 'sync_configured', 'is_default_browser',
           'timezone_cat')

exp_2 <- c('uri_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age', 'is_wow64', 'fxa_configured', 'sync_configured', 'is_default_browser',
           'timezone_offset')

exp_3 <- c('uri_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat', 'is_wow64', 'fxa_configured', 'sync_configured', 'is_default_browser',
           'timezone_cat', 'distro_id_norm')

exp_4 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat', 'is_wow64', 'fxa_configured', 'sync_configured', 'is_default_browser',
           'timezone_cat', 'distro_id_norm') # causing errors when downsampling!, 'os')

exp_5 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'timezone_cat', 'cpu_speed_mhz')

exp_6 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'country', 'cpu_speed_mhz')

exp_7 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'country', 'cpu_speed_mhz', 'memory_mb')

exp_8 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'country', 'cpu_speed_mhz', 'memory_mb', 'cpu_cores')

exp_9 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'country', 'cpu_speed_cat', 'memory_cat', 'cpu_cores_cat')

exp_10 <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 'num_bookmarks',
           'profile_age_cat','fxa_configured', 'sync_configured', 'is_default_browser',
           'country', 'cpu_speed_cat', 'memory_cat', 'cpu_cores_cat', 'distro_id_norm')

exps <- list(exp_1 = exp_1, exp_2 = exp_2, exp_3 = exp_3, exp_4 = exp_4, exp_5 = exp_5, 
             exp_6 = exp_6, exp_7 = exp_7, exp_8 = exp_8, exp_9 = exp_9, exp_10 = exp_10)
```

## Modeling

### CEM: Small Dataset

#### 1x Oversampling
```{r}
cem_results_1x <- list()
cem_models_1x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_1x_sm, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_1x[[exp]] <- res
  cem_models_1x[[exp]] <- cem
  print(res)
}
```

#### 2x Oversampling
```{r}
cem_results_2x <- list()
cem_models_2x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_2x_sm, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_2x[[exp]] <- res
  cem_models_2x[[exp]] <- cem
  print(res)
}
```

#### 4x Oversampling
```{r}
cem_results_4x <- list()
cem_models_4x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_4x_sm, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_4x[[exp]] <- res
  cem_models_4x[[exp]] <- cem
  print(res)
}
```

```{r}
save.image('data/4week/US/analysis_09072019.RData')
```

**Observation**: Adding in environment metrics greatly helps!

### Nearest Neighbors: Small

#### 1x Oversampling
```{r}
nearest_results_1x <- list()
nearest_models_1x <- list()
for (exp in names(exps)){
  nearest <- matchit(formula = generate_formula(exps[[exp]], label), df_1x_sm, 'nearest', replace = TRUE)
  res <- match.data(nearest) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  nearest_results_1x[[exp]] <- res
  nearest_models_1x[[exp]] <- nearest
  print(res)
}
```

#### 2x Oversampling
```{r}
nearest_results_2x <- list()
nearest_models_2x <- list()
for (exp in names(exps)){
  nearest <- matchit(formula = generate_formula(exps[[exp]], label), df_2x_sm, 'nearest')
  res <- match.data(nearest) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  nearest_results_2x[[exp]] <- res
  nearest_models_2x[[exp]] <- nearest
  print(res)
}
```

#### 4x Oversampling
```{r}
nearest_results_4x <- list()
nearest_models_4x <- list()
for (exp in names(exps)){
  nearest <- matchit(formula = generate_formula(exps[[exp]], label), df_4x_sm, 'nearest')
  res <- match.data(nearest) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  nearest_results_4x[[exp]] <- res
  nearest_models_4x[[exp]] <- nearest
  print(res)
}
```

```{r, echo=FALSE}
save.image('data/4week/US/analysis_09072019.RData')
```

### CEM: Full

#### 1x Oversampling
```{r}
cem_results_1x <- list()
cem_models_1x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_1x, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_1x[[exp]] <- res
  cem_models_1x[[exp]] <- cem
  print(res)
}
```

#### 2x Oversampling
```{r}
cem_results_2x <- list()
cem_models_2x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_2x, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_2x[[exp]] <- res
  cem_models_2x[[exp]] <- cem
  print(res)
}
```

#### 4x Oversampling
```{r}
cem_results_4x <- list()
cem_models_4x <- list()
for (exp in names(exps)){
  cem <- matchit(formula = generate_formula(exps[[exp]], label), df_4x, 'cem')
  res <- match.data(cem) %>%
    calc_delta(tr_cov = exps[[exp]], ho_cov = holdout_covariates)
  cem_results_4x[[exp]] <- res
  cem_models_4x[[exp]] <- cem
  print(res)
}
```

### Nearest Neighbors: Small - mahalanobis

Appears `distance = 'mahalanobis'` doesn't work with categorical (singular matrix). Bypassing issue by only using numeric covariates.

#### 1x Oversampling
```{r}
formula <- generate_formula(c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 
                                        'num_bookmarks', 'profile_age', 'timezone_offset', 'cpu_speed_mhz', 'memory_mb', 
                                        'cpu_cores'), label = label)

exp_1x.nearest.ma<- matchit(formula, df_1x_sm, 'nearest', distance='mahalanobis', replace = TRUE)

exp_1x.res.nearest.ma <- match.data(exp_1x.nearest.ma) %>%
  calc_delta(tr_cov = exp_1, ho_cov = holdout_covariates)
exp_1x.res.nearest.ma
```

#### 2x Oversampling
```{r}
exp_2x.nearest.ma.1x<- matchit(formula, df_2x_sm, 'nearest', distance='mahalanobis')

exp_2x.res.nearest.ma <- match.data(exp_2x.nearest.ma) %>%
  calc_delta(tr_cov = exp_1, ho_cov = holdout_covariates)
exp_2x.res.nearest.ma
```

#### 4x Oversampling
```{r}
exp_4x.nearest.ma<- matchit(formula, df_4x_sm, 'nearest', distance='mahalanobis')

exp_4x.res.nearest.ma <- match.data(exp_4x.nearest.ma) %>%
  calc_delta(tr_cov = exp_1, ho_cov = holdout_covariates)
exp_4x.res.nearest.ma
```

```{r, echo=FALSE} 
save.image('data/4week/US/analysis_09072019.RData')
```

```{r}
stats <- calc_stats(df_4x_sm, holdout_covariates, add_1 = TRUE)

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]

  compare_log_cont(df_4x_sm, covariate, means, medians) 
}
```

```{r}
exp_4x.nearest.ma.df <- match.data(exp_4x.nearest.ma)
stats <- calc_stats(exp_4x.nearest.ma.df, holdout_covariates, add_1 = TRUE)

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]

  compare_log_cont(exp_4x.nearest.ma.df, covariate, means, medians) 
}
```

#### 4x: Kitchen Sink
All continuous covariates

```{r}
formula <- generate_formula(c(cont_log_covariates, cont_covariates), label = label)

exp_4x_all.nearest.ma<- matchit(formula, df_4x_sm, 'nearest', distance='mahalanobis')

exp_4x_all.res.nearest.ma <- match.data(exp_4x_all.nearest.ma) %>%
  calc_delta(tr_cov = exp_1, ho_cov = holdout_covariates)
exp_4x_all.res.nearest.ma
```

```{r}
save.image('data/4week/US/analysis_09072019.RData')
```

```{r}
stats <- calc_stats(df_4x_sm, holdout_covariates, add_1 = TRUE)

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]

  compare_log_cont(df_4x_sm, covariate, means, medians) 
}
```

```{r}
exp_4x_all.nearest.ma.df <- match.data(exp_4x_all.nearest.ma)
stats <- calc_stats(exp_4x_all.nearest.ma.df, holdout_covariates, add_1 = TRUE)

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]

  compare_log_cont(exp_4x_all.nearest.ma.df, covariate, means, medians) 
}
```

```{r}
save.image('data/4week/US/analysis_09072019.RData')
```