### 2019-10-01 poc/method_validation.Rmd
Moving reports over

---
title: 'Beta to Release Matching: Method Validation'
author: "Corey Dow-Hygelund"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r imports, warning=FALSE, message=FALSE, echo=FALSE}
source('supporting_funcs.R')
library(kableExtra)
library(cowplot)
```

# Technique

* Apply same data cleaning, covariate/feature generation, and sampling as in [proof-of-concept](https://metrics.mozilla.com/protected/cdowhygelund/beta_subset_release.html)
    - Apply to v67 and v68 datasets
    - **NOTE** Should we sample clients from both versions? 
* Train model (apply matching) to v67
    - Review diagnostics
* Validated against v68
    1. Find all clients in v67 sample 
    2. Review diagnostics

# Data Loading

Follow the same protocol in the [previous work](https://metrics.mozilla.com/protected/cdowhygelund/modeling_poc.Rmd). For this case v67 data is used for modeling, and v68 is used for validation.

```{r}
# define response field
label <- 'is_release'

load_df <- function(file_path){
  df <- read.csv(file_path) %>%
  select(-c(CONTENT_PAINT_TIME_PARENT, COMPOSITE_TIME_CONTENT, COMPOSITE_TIME_PARENT, CONTENT_FRAME_TIME_PARENT)) %>% # All NAs 
  select(-e10s_enabled) %>% # All True
  mutate(default_search_engine = as.factor(normalize_search_engine(default_search_engine))) %>%
  mutate(profile_age_cat = as.factor(normalize_profile_age(profile_age))) %>%
  mutate(distro_id_norm = as.factor(normalize_distro_id(distribution_id))) %>%
  mutate(timezone_cat = as.factor(normalize_timezone(timezone_offset))) %>%
  mutate(memory_cat = as.factor(normalize_memory(memory_mb))) %>% 
  mutate(cpu_speed_cat = as.factor(normalize_cpu_speed(cpu_speed_mhz))) %>% 
  mutate(cpu_cores_cat = as.factor(normalize_cpu_cores(cpu_cores))) %>% 
  mutate(!!label := case_when(
    label == 'beta' ~ FALSE,
    TRUE ~ TRUE)) %>%
  filter(startup_ms >= 0 & profile_age >= 0)

  # complete cases (specifically targets perf metrics) and remove profiles in both Beta and Release
  df_c <- df %>% 
    na.omit() %>%
    mutate()
  return(df_c)
}

df_train <- load_df('data/Validation/df_67.csv')
df_validate <- load_df('data/Validation/df_68.csv')
```

# Sampling
Oversampling Beta by 4x and downsampling this result to 100K records.

```{r}
build_df <- function(df_c, multiple){
  df_beta <- df_c %>% filter(is_release == FALSE)
  df_rel <- df_c %>% filter(is_release == TRUE)
  n_beta <- nrow(df_beta)

  df <- df_rel %>%
    sample_n(size = round(n_beta / multiple)) %>%
    rbind(df_beta)

  return(df)
}

df_train_4x <- build_df(df_train, 4) %>% sample_n(size = 100000)
# df_validate_4x <- build_df(df_validate, 4) %>% sample_n(size = 100000)
```

Determine the number of training (v67) Beta and Release clients that are in the validation set (v68)

```{r}
mutual_clients <- df_train_4x[, c('client_id', 'label')] %>% 
  inner_join(df_validate[, c('client_id', 'label')]) %>% 
  count(label)

mutual_clients
```

Compare this to existing distribution

```{r}
df_train_4x %>% 
  count(label)
```

Okay, not bad most training clients (~70%) are in validation. See how this holds after matching.

# Modeling

```{r covariate_groups}
model_covs <- c('search_count', 'daily_max_tabs', 'daily_num_sessions_started', 'num_addons', 
                              'num_bookmarks', 'profile_age', 'timezone_offset', 'cpu_speed_mhz',
                              'memory_mb', 'cpu_cores')

holdout_covariates <- c("CONTENT_PAINT_TIME_CONTENT", "TIME_TO_DOM_CONTENT_LOADED_END_MS", 
                        "MEMORY_TOTAL", "TIME_TO_DOM_COMPLETE_MS", "FX_PAGE_LOAD_MS_2_PARENT", 
                        "FX_TAB_SWITCH_TOTAL_E10S_MS",
                        "CONTENT_FRAME_TIME_GPU", "COMPOSITE_TIME_GPU",
                        "TIME_TO_LOAD_EVENT_END_MS", 'startup_ms', 'content_crashes')

cont_log_covariates <- c('active_hours', 'daily_max_tabs', 'daily_tabs_opened', 'search_count',
                         'daily_unique_domains', 'daily_num_sessions_started', 
                         'num_bookmarks', 'num_addons', 'num_active_days', 'num_pages', 'uri_count',
                         'session_length', 'profile_age')

cat_covariates <- c('country', 'timezone_cat', 'distro_id_norm', 'profile_age_cat', 'cpu_cores_cat',
                    'memory_cat', 'cpu_speed_cat', 'default_search_engine')

```

The best model from previous work:

```{r model_train}
formula <- generate_formula(model_covs, label = label)

nn.ma <- matchit(formula, df_train_4x, 'nearest', distance='mahalanobis')

df_matched <- match.data(nn.ma) 
```
```{r export_model, warning=FALSE, message=FALSE, echo=FALSE}
save.image('data/Validation/analysis_09182019_corr.RData')
```

```{r summary_mean_tbl}
stats_pre <- calc_delta(df_train_4x, NULL, holdout_covariates) %>%
  select(-'content_crashes')
stats_post <- calc_delta(df_matched, NULL, holdout_covariates) %>%
  select(-'content_crashes')

stats_mean <- stats_pre[1, ] %>% 
  rbind(stats_post[1, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

stats_median <- stats_pre[2, ] %>% 
  rbind(stats_post[2, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

knitr::kable(stats_mean) %>%
  kable_styling() %>%
  scroll_box(width = "750px", height = "200px")

```

```{r perf_holdout, fig.width=10, fig.height=50, echo=FALSE}
df_full <- df_matched %>% 
  select(-c(distance, weights)) %>%
  filter(label == 'beta') %>%
  mutate(label =  'beta - matched') %>%
  rbind(df_train)

stats <- calc_stats(df_matched, holdout_covariates, add_1 = TRUE)

plots <- list()

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]
  ho_means <- stats %>% filter(label == 'beta' & metric == 'mean') %>% select(covariate)

  plots[[covariate]] <- compare_log_cont(df_full, covariate, means, medians, as.numeric(ho_means), print=FALSE) 
}

plot_grid(plotlist = plots, ncol = 1)

# for (covariate in holdout_covariates) {
#   compare_log_cont(df_full, covariate) 
#}
```

# Validation

Subset the validation clients down to those matched

```{r}
df_validate_matched <- df_validate %>%
  filter(client_id %in% df_matched$client_id)

df_validate_matched %>%
  count(label)
```

## Holdout Covariates

```{r summary_validate_mean_tbl}
stats_pre <- calc_delta(df_validate, NULL, holdout_covariates) %>%
  select(-'content_crashes')
stats_post <- calc_delta(df_validate_matched, NULL, holdout_covariates) %>%
  select(-'content_crashes')

stats_mean <- stats_pre[1, ] %>% 
  rbind(stats_post[1, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

stats_median <- stats_pre[2, ] %>% 
  rbind(stats_post[2, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

knitr::kable(stats_mean) %>%
  kable_styling() %>%
  scroll_box(width = "750px", height = "200px")

```

```{r}
knitr::kable(stats_median) %>%
  kable_styling() %>%
  scroll_box(width = "750px", height = "200px")
```

```{r perf_holdout, fig.width=10, fig.height=50, echo=FALSE}
df_validate_full <- df_validate_matched %>% 
  # select(-c(distance, weights)) %>%
  filter(label == 'beta') %>%
  mutate(label =  'beta - matched') %>%
  rbind(df_validate) %>% 
  distinct()

stats <- calc_stats(df_validate_full, holdout_covariates, add_1 = TRUE)

plots <- list()

for (covariate in holdout_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]
  ho_means <- stats %>% filter(label == 'beta' & metric == 'mean') %>% select(covariate)

  plots[[covariate]] <- compare_log_cont(df_validate_full, covariate, means, medians, as.numeric(ho_means), print=FALSE) 
}

# for (covariate in holdout_covariates) {
#   compare_log_cont(df_full, covariate) 
#}
```

```{r perf_holdout, fig.width=10, fig.height=50, echo=FALSE}
plot_grid(plotlist = plots, ncol = 1)
```

## Training/Engagement Covariates

```{r summary_validate_mean_tbl}
stats_pre <- calc_delta(df_validate, NULL, cont_log_covariates) 
stats_post <- calc_delta(df_validate_matched, NULL, cont_log_covariates)

stats_mean <- stats_pre[1, ] %>% 
  rbind(stats_post[1, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

stats_median <- stats_pre[2, ] %>% 
  rbind(stats_post[2, ]) %>%
  set_rownames(c('pre-matching', 'post-matching'))

knitr::kable(stats_mean) %>%
  kable_styling() %>%
  scroll_box(width = "750px", height = "200px")

```

```{r}
knitr::kable(stats_median) %>%
  kable_styling() %>%
  scroll_box(width = "750px", height = "200px")
```

```{r cont_log_cov_plots, fig.width=10, fig.height=50, echo=FALSE}
stats <- calc_stats(df_validate_full, cont_log_covariates, add_1 = TRUE)

plots <- list()

for (covariate in cont_log_covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]
  ho_means <- stats %>% filter(label == 'beta' & metric == 'mean') %>% select(covariate)

  plots[[covariate]] <- compare_log_cont(df_validate_full, covariate, means, medians, as.numeric(ho_means), print=FALSE) 
}
```

```{r cont_log_cov_plotted, fig.width=10, fig.height=50, echo=FALSE}
plot_grid(plotlist = plots, ncol = 1)
```

Categorical

```{r cat_cov_plotted, fig.width=10, fig.height=50, echo=FALSE}
plots <- list()

for (covariate in cat_covariates){
  plots[[covariate]] <- compare_cat(df_validate_full, covariate, limit = 10, plot = FALSE)
}

plot_grid(plotlist = plots, ncol = 1)
```