### 2019-10-15 perf_release_criteria/scoring_algo_analysis.Rmd
Adding in testing out scoring methods. Cramer von Mises it is.

---
title: 'Milestone 2 Test: Scoring Algorithms'
output:
  html_document:
    df_print: paged
---

# tl;dr 
This is analysis to determine the feasibility of the various [scoring algorithms](https://docs.google.com/document/d/1SfuanvmYmvmEFAdQ7Z5djDeLezdNB1TESqVmj93O8to/edit#heading=h.w3huje1nj766) proposed in the modeling approach for this [PRD](https://docs.google.com/document/d/1Ygz6MkudYHZjnDnD9Z97kUyFrvV3KGWsjXyPjddhHq0/edit#heading=h.lvb9l8gw2nee).

```{r imports, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(transport)
library(profmem)
```

# Method
* Utilize previous saved image from proof-of-concept validation
* Compare performance metrics of Beta to Release, without any matching. On test. 
* Benchmark computation time, and memory usage (if necessary)

```{r data_load, echo=FALSE}
load('data/milestone2/analysis_10022019_corr.RData')
```

# transport
1st, try out the `transport` library, to determine if Wasserstein distance will work with out dataset.

## Wasserstein
The relevant `transport` class appears to be a `pp`, or "discrete measures with some fixed mass at any of finitely many locations". Basically each instance corresponds to a mass = 1, and the corresponding metric values are coordinates to where that point is located.

1st, normalize each performance distributions to unit length. This puts them all on the same scale, so that the "cost" of transport is equal between the dimensions.

```{r}
perf_metrics <- c("TIME_TO_DOM_CONTENT_LOADED_END_MS","TIME_TO_DOM_COMPLETE_MS", "FX_PAGE_LOAD_MS_2_PARENT", "TIME_TO_LOAD_EVENT_END_MS")

norm <- function(x) return(x/sum(x))

df_norm <- df_train %>%
  select(perf_metrics, label) %>%
  na.omit() %>%
  mutate_at(vars(perf_metrics), norm)

beta <- df_norm %>% 
  filter(label == 'beta') %>%
  select(perf_metrics)
  # na.omit() #  %>%
  # mutate_all(norm)

release <- df_norm %>% 
  filter(label == 'release') %>% 
  select(perf_metrics)
  # na.omit() # %>%
  # mutate_all(norm)
```

Calculate the Wasserstein distance. It appears that the same number of instances must be in both distributions.
**NOTE**: Received `Error in gen_cost(x, y, 1) : std::bad_alloc` when using 85K records. Downsample to 30K.

```{r wassertstein, error=TRUE}
wd <- wasserstein(pp(beta %>% sample_n(30000)), pp(release %>% sample_n(30000)))
```

>>> Error: cannot allocate vector of size 6.7 Gb

Okay this isn't going to work. Reduce down to single dimension for the earth mover's distance. Use `FX_PAGE_LOAD_MS_2`. Utilize the original dataset values. Benchmark the time.

## EMD

### Pre-Matching
```{r emd}
beta <- df_train %>% 
  filter(label == 'beta') %>%
  select(perf_metrics) %>%
  na.omit()

release <- df_train %>% 
  filter(label == 'release') %>%
  select(perf_metrics) %>%
  na.omit()

system.time(emd <- wasserstein1d(beta$FX_PAGE_LOAD_MS_2_PARENT, release$FX_PAGE_LOAD_MS_2_PARENT))
print(emd)
```

That is reasonable for the use-case: ~10s for full dataset and all metrics. See memory usage:

```{r emd_mem}
emd_mem <- profmem(emd <- wasserstein1d(beta$FX_PAGE_LOAD_MS_2_PARENT, release$FX_PAGE_LOAD_MS_2_PARENT, p=2))
total(emd_mem)/(1024^2)
```

### Post-matching

```{r emd_matched}
beta_m <- df_matched %>% 
  filter(label == 'beta') %>%
  select(perf_metrics) %>%
  na.omit()

release_m <- df_matched %>% 
  filter(label == 'release') %>%
  select(perf_metrics) %>%
  na.omit()

system.time(emd <- wasserstein1d(beta_m$FX_PAGE_LOAD_MS_2_PARENT, release_m$FX_PAGE_LOAD_MS_2_PARENT))
print(emd)
```

## Cramér-von Mises
Try a Cramér-von Mises (`p=2`).

```{r cms}
system.time(cmd <- wasserstein1d(beta$FX_PAGE_LOAD_MS_2_PARENT, release$FX_PAGE_LOAD_MS_2_PARENT, p=2))
print(cmd)
```

This works also! Onto memory profiling.

```{r cms_mem}
cms_mem <- profmem(cmd <- wasserstein1d(beta$FX_PAGE_LOAD_MS_2_PARENT, release$FX_PAGE_LOAD_MS_2_PARENT, p=2))
total(cms_mem)/(1024^2)
```

### Post-matching

```{r cmd_matched}
system.time(cmd <- wasserstein1d(beta_m$FX_PAGE_LOAD_MS_2_PARENT, release_m$FX_PAGE_LOAD_MS_2_PARENT, p=2))
print(cmd)
```