### 2019-10-25 perf_release_criteria/milestone_3/feature_selection_nn_logit_linear.Rmd
Completed analysis for nn logit linear.

---
title: "Milestone 2: Feature Selection - Nearest Neighbors, Logit w/Linear Propensity Score"
author: "Corey Dow-Hygelund, Mozilla Data Science"
date: 'Last Updated: `r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_notebook:
    theme: cosmo
    toc: true
    toc_float: true
---

# tl;dr

This analysis determines the relevant [covariates (features)](https://docs.google.com/document/d/1SfuanvmYmvmEFAdQ7Z5djDeLezdNB1TESqVmj93O8to/edit#heading=h.nwwrif453n80) to utilize for computationally intensive statistical matching algorithms proposed in the modeling approach for this [PRD](https://docs.google.com/document/d/1Ygz6MkudYHZjnDnD9Z97kUyFrvV3KGWsjXyPjddhHq0/edit#heading=h.lvb9l8gw2nee).

**Matching Algorithm Tested**: Nearest-neighbors with logit distance metric transformed back to linear scale.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source('../lib/supporting_funcs.R')
source('../lib/scoring.R')
library(MatchIt)
```

```{r data_load, echo=FALSE}
file_name = 'df_train_validate_20191021.RData'
image_file_path = file.path('data', file_name)

# Pull from GCP if necessary
# library(googleCloudStorageR)
# Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
#            "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
# gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)

load(image_file_path)
```

```{r bts, echo=FALSE}
bts = list(
  bts_1x = bts.1x,
  bts_4x = bts.4x,
  bts_8x = bts.8x,
  bts_16x = bts.16x
)

perf_metrics <- names(get_m2_metric_map())

covs <- df_train_f %>%
  select(-perf_metrics) %>%
  select(-content_crashes) %>%
  select(-client_id) %>%
  select(-label) %>%
  select(-is_release) %>%
  select(-app_version) %>%
  # select_if(is.numeric) %>% # Mahalanobis constraint
  names()

results_file_name <- 'feature_selection_nn_logit_linear.RData' 
results_file_path <- file.path('data', results_file_name)
```

# Features
Feature sets:

1. Only performance metrics.
2. Performance metrics and the highest found by Boruta with original dataset.
3. Performance metrics and the highest found by Boruta with equal labels dataset.
4. Performance metrics and all covariates.
5. Only the highest covariates found by Boruta (excluding performance).
6. Only the highest covariates found by Boruta (equal labels, excluding performance).
7. Utilize all covariates (excluding performance).

Retrieve Boruta features:

```{r boruta_import}
file_name = 'feature_selection_boruta_initial_20191023.RData'
image_file_path = file.path('data', file_name)

Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
library(googleCloudStorageR)
gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)

load(image_file_path)
```

```{r boruta_fs}
extract_boruta_fs <- function(boruta_res, num=5){
  features <- NULL
  for(metric in names(boruta_results)){
    features <- c(names(sort(apply(boruta_res[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:num]), features)
  }
  return(sort(unique(features)))
}

features_top10 <- extract_boruta_fs(boruta_results, num=10)
features_top10_eq <- extract_boruta_fs(boruta_results_eq, num=10)

# filter out categorical
features_top10 <- df_train_f %>% 
  select(features_top10) %>% 
  # select_if(is.numeric) %>% 
  names()
features_top10_eq <- df_train_f %>% 
  select(features_top10_eq) %>% 
  # select_if(is.numeric) %>% 
  names()
```

```{r feature_sets}
fs1 <- perf_metrics
fs2 <- c(names(fs1), features_top10)
fs3 <- c(names(fs1), features_top10_eq)
fs4 <- c(names(fs1), covs)
fs5 <- features_top10
fs6 <- features_top10_eq
fs7 <- covs
```

# FS 1
Feature selection on only performance metrics

```{r fs1, error=TRUE}
res_fs1 <- list()
for (bt in names(bts)){
  res_fs1[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs1, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs1)
```

```{r fs1_noreplace, error=TRUE}
res_fs1_noreplace <- list()
for (bt in names(bts)){
  res_fs1_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs1, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs1_noreplace)
```

# FS 2

Performance metrics and the highest found by Boruta with original dataset.

```{r fs2, error=TRUE}
res_fs2 <- list()
for (bt in names(bts)){
  res_fs2[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs2, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs2)
```

```{r fs2_noreplace, error=TRUE}
res_fs2_noreplace <- list()
for (bt in names(bts)){
  res_fs2_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs2, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs2_noreplace)
```

# FS 3

Performance metrics and the highest found by Boruta with equal labels dataset.

```{r fs3, error=TRUE}
res_fs3 <- list()
for (bt in names(bts)){
  res_fs3[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs3, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs3)
```

```{r fs3_noreplace, error=TRUE}
res_fs3_noreplace <- list()
for (bt in names(bts)){
  res_fs3_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs3, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs3_noreplace)
```

# FS 4

Performance metrics and all covariates.

```{r fs4, error=TRUE}
res_fs4 <- list()
for (bt in names(bts)){
  res_fs4[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs4, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs4)
```

```{r fs4_noreplace, error=TRUE}
res_fs4_noreplace <- list()
for (bt in names(bts)){
  res_fs4_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs4, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs4_noreplace)
```

# FS 5

Only the highest covariates found by Boruta (excluding performance).

```{r fs5, error=TRUE}
res_fs5 <- list()
for (bt in names(bts)){
  res_fs5[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs5, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs5)
```

```{r fs5_noreplace, error=TRUE}
res_fs5_noreplace <- list()
for (bt in names(bts)){
  res_fs5_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs5, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs5_noreplace)
```

# FS 6

Only the highest covariates found by Boruta (equal labels, excluding performance).

```{r fs6, error=TRUE}
res_fs6 <- list()
for (bt in names(bts)){
  res_fs6[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs6, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs6)
```

```{r fs6_noreplace, error=TRUE}
res_fs6_noreplace <- list()
for (bt in names(bts)){
  res_fs6_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs6, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs6_noreplace)
```

# FS 7

Utilize all covariates (excluding performance).

```{r fs7, error=TRUE}
res_fs7 <- list()
for (bt in names(bts)){
  res_fs7[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs7, size=NULL, replace=TRUE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs7)
```

```{r fs7_noreplace, error=TRUE}
res_fs7_noreplace <- list()
for (bt in names(bts)){
  res_fs7_noreplace[[bt]] <- perform_matchit_fs(df_train_f, bts[[bt]], fs7, size=NULL, replace=FALSE, workers = 10, distance = 'linear.logit')
}
save.image(results_file_path)
print(res_fs7_noreplace)
```

# Serialize and Push to GCP
Save the resultant dataset and bootstrap samples to an R image.

```{r serialize_dataset}
save(list = ls(pattern = 'res_fs'), file = results_file_path)
```

```{r gcp_storage_push, warning=FALSE}
gcs_file_path <- file.path('data', 'milestone2', results_file_name)

Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")

library(googleCloudStorageR)

upload_try <- gcs_upload(results_file_path, name = gcs_file_path)
upload_try
```