### 2019-11-06 perf_release_criteria/milestone_3/ms_2_validation.Rmd
Adding in text.

---
title: 'Milestone 2: Model Validation'
output:
  html_notebook:
    theme: cosmo
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
date: 'Last Updated: `r format(Sys.time(), "%B %d, %Y")`'
---

# tl;dr 
This notebook determines the model framework utilized for this [PRD](https://docs.google.com/document/d/1Ygz6MkudYHZjnDnD9Z97kUyFrvV3KGWsjXyPjddhHq0/edit#heading=h.lvb9l8gw2nee). It utilizes the results from the hyperparameter tuning step, and trains the optimal model for each algorithm on the v67 dataset. The resultant predictions of these optimal models are compared to a v68 validation dataset. The framework used to generate the highest performant model (e.g., features, hyperparameters) will be used in online model training for the subsetting of Beta in production use.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source('../lib/supporting_funcs.R')
source('../lib/scoring.R')
library(MatchIt)
```

```{r data_load, echo=FALSE}
file_name = 'df_train_validate_20191025.RData'
image_file_path = file.path('data', file_name)

# Pull from GCP if necessary
if (!file.exists(image_file_path)){
  Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
  library(googleCloudStorageR)
  gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)
}

load(image_file_path)
```

```{r var_def, echo=FALSE}
df_rel_val <- df_validate_f %>%
  filter(label == 'release')

df_beta_train <- df_train_f %>% filter(is_release == FALSE)
df_rel_train <- df_train_f %>% filter(is_release == TRUE)
n_beta <- nrow(df_beta_train)
```

# Features
Feature sets:

1. Only performance metrics.
2. Performance metrics and the highest found by Boruta with original dataset.
3. Performance metrics and the highest found by Boruta with equal labels dataset.
4. Performance metrics and all covariates.
5. Only the highest covariates found by Boruta (excluding performance).
6. Only the highest covariates found by Boruta (equal labels, excluding performance).
7. Utilize all covariates (excluding performance).

Retrieve Boruta features:

```{r boruta_import, echo=FALSE, warning=FALSE, message=FALSE}
file_name = 'feature_selection_boruta_initial_20191023.RData'
image_file_path = file.path('data', file_name)

if (!file.exists(image_file_path)){
  Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
  library(googleCloudStorageR)
  gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)
}
load(image_file_path)
```

```{r boruta_fs, echo=FALSE}
extract_boruta_fs <- function(boruta_res, num=5){
  features <- NULL
  for(metric in names(boruta_results)){
    features <- c(names(sort(apply(boruta_res[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:num]), features)
  }
  return(sort(unique(features)))
}

features_top10 <- extract_boruta_fs(boruta_results, num=10)
features_top10_eq <- extract_boruta_fs(boruta_results_eq, num=10)

# filter out categorical
features_top10 <- df_train_f %>% 
  select(features_top10) %>% 
  select_if(is.numeric) %>% 
  names()
features_top10_eq <- df_train_f %>% 
  select(features_top10_eq) %>% 
  select_if(is.numeric) %>% 
  names()
```

```{r feature_sets, echo=FALSE}
perf_metrics <- names(get_m2_metric_map())

covs <- df_train_f %>%
  select(-perf_metrics) %>%
  select(-content_crashes) %>%
  select(-client_id) %>%
  select(-label) %>%
  select(-is_release) %>%
  select(-app_version) %>%
  select_if(is.numeric) %>% # Mahalanobis constraint
  names()

fs1 <- perf_metrics
fs2 <- c(names(fs1), features_top10)
fs3 <- c(names(fs1), features_top10_eq)
fs4 <- c(names(fs1), covs)
fs5 <- features_top10
fs6 <- features_top10_eq
fs7 <- covs
```

# Validation Bootstraps

30 bootstrap replicates of v68 Release will be utilized to validate the models. There are `r nrow(df_rel_val)` profiles utilized for this purpose.

```{r bts_validate}
bts = list()
for(i in 1:30){
  bts[[i]] <- df_rel_val %>% 
    sample_frac(size = 1, replace = TRUE) %>%
    pull(client_id) 
}
```

```{r scorer}
score_model <- function(bts, df_match, df_val, workers){
  if (missing(workers)) workers = detectCores()
  cl <- makePSOCKcluster(workers) 
  registerDoParallel(cl)
  final <- tryCatch({
    scores <- foreach(i=1:length(bts), 
                      .packages = c('dplyr', 'transport'), 
                      .export=c('calc_score', 'calc_cms', 'get_m2_metric_map')) %dopar% {
                        bt <- bts[[i]]
                        test <- df_val %>% 
                          right_join(data.frame(client_id = bt, stringsAsFactors=FALSE), by='client_id', 'right')

                        df_scores <- test %>%
                          bind_rows(df_match)

                        score <- calc_score(df_scores, get_m2_metric_map())
                        score
                        }
    scores <- unlist(scores)
    c(mean = mean(scores), median = median(scores))
  }, 
  error = function(cond){
    message(paste("Bootstrap validation failed: ", cond))
    return(NA)
  },
  finally = {
    stopCluster(cl)
  }
  )
  return(final)
}
```

# Oversampling

The entire v67 dataset will be utilized for training. However, most of the matching algorithms are more performant with higher ratios of Beta to Release. There are `r n_beta` profiles utilized for this purpose.

```{r oversampling}
oversample <- function(oversampling, df_beta, df_rel) {
  df_x <- df_rel %>%
    sample_n(size = round(n_beta / oversampling)) %>%
    rbind(df_beta)
  return(df_x)
  }

oversamples <- c(1, 2, 4, 8, 16)
dfs <- lapply(oversamples, oversample, df_beta = df_beta_train, df_rel = df_rel_train)
names(dfs) <- as.character(oversamples)
```

# Model Training

Two matching methods were ultimately tested, as CEM and subclassing were producing extremely poor results.

* nearest-neighbors
* genetic matching

Feature selection and hyperparameter tuning for the nearest-neighbors models were all trained using the `MatchIt` library.  The following method trains these models using the full training dataset.

```{r trainer}
train_matchit <- function(train, model_covs, add_interactions, ...){
  # train model
  formula <- generate_formula(model_covs, label = 'is_release', add_interactions)
  model <- matchit(formula, train, ...)

  # extract beta subset
  df_matched <- get_matches(model, train) %>%
    select(-weights, -distance) %>%
    filter(label == 'beta')

  return(list(model = model, matched = df_matched))
}

```

## Nearest-Neighbors, Malahanobis

Hyperparameters: `hyperparameter_tuning_nn_malahanobis.Rmd`

* `caliper` = 0.25
* `calcloset` = TRUE
* `ratio` = 2
* `replace` = FALSE

```{r nn_mal_fs1}
nn.mal.fs1 <- train_matchit(dfs[['4']], fs1, add_interactions = FALSE, replace = FALSE, 
                        caliper = 0.25, calclosest = TRUE, ratio = 2, distance = "mahalanobis")
score_model(bts, nn.mal.fs1$matched, df_rel_val)
```

```{r nn_mal_fs3}
nn.mal.fs3 <- train_matchit(dfs[['16']], fs3, add_interactions = FALSE, replace = TRUE, 
                        caliper = 0.25, calclosest = TRUE, ratio = 1, distance = "mahalanobis")
score_model(bts, nn.mal.fs3$matched, df_rel_val)
```

```{r nn_mal_fs5}
nn.mal.fs5 <- train_matchit(dfs[['8']], fs5, add_interactions = FALSE, replace = TRUE, 
                        caliper = 0.25, calclosest = TRUE, ratio = 1, 
                        distance = "mahalanobis")
score_model(bts, nn.mal.fs3$matched, df_rel_val)
```

## GAM

# TODO

* Build validation methods
    - Plots
    - Scoring
    - Everything I would use for dashboard

# Serialize

* The bootstraps and training datasets!

```{r}
save(nn.mal, 
     file = '')
```