### 2019-11-06 perf_release_criteria/milestone_3/hyperparameter_tuning/hyperparameter_tuning_nn_logit_linear.Rmd

---
title: "Milestone 2: Hyperparameter Tuning - Nearest Neighbors, Logit w/Linear Propensity Score"
author: "Corey Dow-Hygelund, Mozilla Data Science"
date: 'Last Updated: `r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_notebook:
    theme: cosmo
    toc: true
    toc_float: true
---

# tl;dr

This analysis determines the optimal [hyperparameters](https://docs.google.com/document/d/1SfuanvmYmvmEFAdQ7Z5djDeLezdNB1TESqVmj93O8to/edit#heading=h.yqukkyb8q4xw) to utilize for computationally intensive statistical matching algorithms proposed in the modeling approach for this [PRD](https://docs.google.com/document/d/1Ygz6MkudYHZjnDnD9Z97kUyFrvV3KGWsjXyPjddhHq0/edit#heading=h.lvb9l8gw2nee).

**Matching Algorithm Tested**: Nearest-neighbors with logit distance metric transformed back to linear scale.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source('../lib/supporting_funcs.R')
source('../lib/scoring.R')
library(MatchIt)
```

```{r data_load, echo=FALSE}
file_name = 'df_train_validate_20191025.RData'
image_file_path = file.path('data', file_name)

# Pull from GCP if necessary
if (!file.exists(image_file_path)){
  Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
  library(googleCloudStorageR)
  gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)

}
load(image_file_path)
```

```{r bts, echo=FALSE}
bts = list(
  bts_1x = bts.1x,
  bts_4x = bts.4x,
  bts_8x = bts.8x,
  bts_16x = bts.16x
)

perf_metrics <- names(get_m2_metric_map())

covs <- df_train_f %>%
  select(-perf_metrics) %>%
  select(-content_crashes) %>%
  select(-client_id) %>%
  select(-label) %>%
  select(-is_release) %>%
  select(-app_version) %>%
  names()

results_file_name <- 'hyperparameter_tuning_nn_logit_linear.RData' 
results_file_path <- file.path('data', results_file_name)
```

# Features
The optimal feature set was found in `feature_selection_nn_logit_linear.Rmd`. At the time of this run they were:

* Beta oversampling by 4x
* FS 1 Interactions: Only performance metrics, with interaction terms

To investigate, and possible utilize, a more general model, the second highest suite of features will also be tested:

* Beta oversapmling by 8x
* FS 5: Only the highest covariates found by Boruta (excluding performance)

```{r boruta_import, echo=FALSE, warning=FALSE, message=FALSE}
file_name = 'feature_selection_boruta_initial_20191023.RData'
image_file_path = file.path('data', file_name)

# Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
#            "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")
# library(googleCloudStorageR)
# gcs_get_object(file.path('data', 'milestone2', file_name), saveToDisk = image_file_path, overwrite = TRUE)

load(image_file_path)
```

```{r boruta_fs, echo=FALSE, warning=FALSE, message=FALSE}
extract_boruta_fs <- function(boruta_res, num=5){
  features <- NULL
  for(metric in names(boruta_results)){
    features <- c(names(sort(apply(boruta_res[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:num]), features)
  }
  return(sort(unique(features)))
}

features_top10 <- extract_boruta_fs(boruta_results, num=10)
features_top10_eq <- extract_boruta_fs(boruta_results_eq, num=10)

# filter out categorical
features_top10 <- df_train_f %>% select(features_top10) %>% select_if(is.numeric) %>% names()
features_top10_eq <- df_train_f %>% select(features_top10_eq) %>% select_if(is.numeric) %>% names()
```

```{r feature_sets}
fs1 <- perf_metrics
fs2 <- c(names(fs1), features_top10)
fs3 <- c(names(fs1), features_top10_eq)
fs4 <- c(names(fs1), covs)
fs5 <- features_top10
fs6 <- features_top10_eq
fs7 <- covs
```

# Hyperparameters
Hyperparameters tested for nearest-neighbors are the following:

* Matching order (m.order)
* Caliper
* of control to treatments

```{r hp}
hps <- list(
  replace = c(TRUE, FALSE),
  caliper = c(0, 0.25, 0.5, 1, 3),
  calclosest = c(TRUE, FALSE),
  ratio = c(1, 2, 4)
)

hps_grid = expand.grid(hps)
```

# FS 1
## Coarse Grid
Hyperparameter tuning on optimal features.

```{r fs1, error=TRUE}
fs1_results <- perform_matchit_hp_tune(df_train_f, bts[['bts_4x']], fs1, size=NULL, workers = 64, hps_grid = hps_grid, 
                                       distance = 'linear.logit', add_interactions=TRUE)

fs1_results_df <- fs1_results %>% bind_rows()
fs1_results_df %>% group_by_at(setdiff(names(fs1_results_df), "scores")) %>% 
  summarise(mean = mean(scores), sd = sd(scores), median = median(scores)) -> fs1_results_df
knitr::kable(fs1_results_df[order(fs1_results_df$mean), ] %>% head(20))
```

## Fine Grid

```{r hp_fine}
hps_fine <- list(
  replace = c(TRUE, FALSE),
  caliper = seq(0.2, 0.8, by=0.2),
  calclosest = c(TRUE, FALSE),
  ratio = c(1, 2, 3)
)

hps_fine_grid = expand.grid(hps_fine)
```

```{r fs1_fine, error=TRUE}
hps_fine_grid = expand.grid(hps_fine)

fs1_results_final <- perform_matchit_hp_tune(df_train_f, bts[['bts_4x']], fs1, size=NULL, workers = 64, hps_grid = hps_fine_grid, 
                                       distance = 'linear.logit', add_interactions=TRUE)
save(list=ls(pattern = 'fs[0-9]_results'),  file = results_file_path)

fs1_results_df <- fs1_results_final %>% bind_rows()
fs1_results_df %>% group_by_at(setdiff(names(fs1_results_df), "scores")) %>% 
  summarise(mean = mean(scores), sd = sd(scores), median = median(scores)) -> fs1_results_df
knitr::kable(fs1_results_df[order(fs1_results_df$mean), ] %>% head(20))
```

# FS 5

```{r fs5, error=TRUE}
fs3_results <- perform_matchit_hp_tune(df_train_f, bts[['bts_8x']], fs5, size=NULL, workers = 64, hps_grid = hps_grid, 
                                       distance = 'linear.logit', add_interactions=FALSE)
save(list=ls(pattern = 'fs[0-9]_results'),  file = results_file_path)

fs3_results_df <- fs3_results %>% bind_rows()
fs3_results_df %>% group_by_at(setdiff(names(fs3_results_df), "scores")) %>% 
  summarise(mean = mean(scores), sd = sd(scores), median = median(scores)) -> fs3_results_df
knitr::kable(fs3_results_df[order(fs3_results_df$mean), ] %>% head(20))
```

# Serialize and Push to GCP
Save the resultant dataset and bootstrap samples to an R image.

```{r gcp_storage_push, warning=FALSE}
gcs_file_path <- file.path('data', 'milestone2', results_file_name)

Sys.setenv("GCS_DEFAULT_BUCKET" = "moz-fx-dev-subbeta",
           "GCS_AUTH_FILE" = "moz-fx-dev-cdowhyglund-subBeta-788f8f0d4627.json")

library(googleCloudStorageR)

upload_try <- gcs_upload(results_file_path, name = gcs_file_path)
upload_try
```