### 2019-12-05 perf_release_criteria/milestone_4/training_data_pull.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(bigrquery)
```

```{r}
project <- "moz-fx-dev-cdowhyglund-subbeta"
```

```{r}
# This doesn't work yields error: "Error: Don't know how to convert type numeric"

# sql <- "
# SELECT *
# from `moz-fx-data-shared-prod.analysis.cdowhygelund_subBeta_training` 
# "
#  
# train_df <- query_exec(sql, project = project, use_legacy_sql = FALSE)
```

```{r}
library(DBI)
library(dplyr)

con <- dbConnect(
  bigrquery::bigquery(),
  project = 'moz-fx-data-shared-prod',
  dataset = 'analysis',
  billing = project
)

training_tbl <- tbl(con, "cdowhygelund_subBeta_training")

df <- training_tbl %>%
  collect()
```