### 2019-12-06 poc/EDA/EDA.Rmd
creating EDA directory + initializing R markdown

---
title: "Mozilla Data Science Project - 2019"
author: "Mariana de Oliveira Santos Silva"
date: "December 05, 2019"
output: 
  html_document:
    toc: true
    theme: united
    highlight: textmate
---

# Exploratory Data Analysis (EDA) {.tabset}

This R markdown covers the Exploratory Data Analysis (EDA) on the training dataset of the PRD.

## Setting-up

```{r message=FALSE, warning=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
```

```{r message=FALSE, warning=FALSE}
## Loading the training dataset

load("~/GitHub/ff-beta-release-matching/poc/EDA/data_milestone2_df_train_validate_20191021.RData")
```

```{r message=FALSE, warning=FALSE}
## View train dataframe

kable(head(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Step 1 : Data Inspection

To get introduced to our dataset:

```{r}
kable(introduce(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

Let's have a look on the basic information of the dataset.

```{r, fig.width = 10, fig.height = 5}
plot_intro(df_train_f, ggtheme = theme_minimal())
```

Sanity check on the structure and show sample data for each variable.

```{r}
str(df_train_f)
```

When we get a new dataset to analyze, is important to know if there are missing values and the data type. The `df_status` function can help us by showing these numbers in relative and percentage values.

```{r}
kable(df_status(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

* `q_zeros`: quantity of zeros (`p_zeros`: in percent)
* `q_inf`: quantity of infinite values (`p_inf`: in percent)
* `q_na`: quantity of NA (`p_na`: in percent)
* `type`: factor, ordered-factor, numeric, integer or character
* `unique`: quantity of unique values

> Are all the variables in the correct data type?

`to-do`

<!-- Some variables are encoded as numbers, but they are codes or categories and the models don't handle them in the same way. -->

> Any variables with lots of zeros?

Yes. Variables with lots of zeros may not be useful for modeling and, in some cases, they may dramatically bias the model.

> Any variables with lots of NAs?

None. Good news.

> Any high cardinality variable?

`to-do`
<!-- Factor/categorical variables with a high number of different values (~30) tend to do overfitting if the categories have low cardinality. -->

## Step 2 : Analyzing Categorical Variables

```{r}
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

freq(df_train_f$label)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution of all discrete variables
plot_bar(df_train_f)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution for release dataframe
plot_bar(df_release)
```

```{r, fig.width = 15, fig.height = 15}
## Frequency distribution beta dataframe
plot_bar(df_beta)
```

### EDA on Coninuous Variables

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_train_f)
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_release)
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of all continuous variables
plot_histogram(df_beta)
```