### 2019-12-12 poc/EDA/EDA.Rmd
updating the EDA notebook

---
title: "Mozilla Data Science Project - 2019"
author: "Mariana de Oliveira Santos Silva"
date: "December 05, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    df_print: kable
    code_folding: "hide"
    theme: "flatly"
---

# Exploratory Data Analysis (EDA)

This R markdown covers the Exploratory Data Analysis (EDA) on the training dataset of the PRD.

## Setting-up
```{r message=FALSE, warning=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
```

```{r message=FALSE, warning=FALSE}
## Loading the training dataset

load("~/GitHub/ff-beta-release-matching/poc/EDA/data_milestone2_df_train_validate_20191025.RData")
```

## Training

```{r message=FALSE, warning=FALSE}
## View train dataframe

kable(head(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Validation

```{r message=FALSE, warning=FALSE}
## View train dataframe

kable(head(df_validate_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Step 1 : Data Inspection

### Training

To get introduced to our training dataset, let's have a look on the basic information of the dataset.

```{r echo=FALSE}
kable(introduce(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

```{r echo=FALSE, fig.height=10, fig.width=10}
plot_intro(df_train_f, ggtheme = theme_minimal())
```

### Validation

To get introduced to our validation dataset, let's have a look on the basic information of the dataset.

```{r echo=FALSE}
kable(introduce(df_validate_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

```{r echo=FALSE, fig.height=10, fig.width=10}
plot_intro(df_validate_f, ggtheme = theme_minimal())
```

### Observations

* Most of the dataset is composed of continuous variables
* No NAs values (were handled in preprocessing)

## Data Structure

### Training

Let's use `glimpse` function to display a vertical preview of the training dataset. So we can easily preview data type and sample data.

```{r}
glimpse(df_train_f)
```

If we want to get some metrics about data types, zeros, infinite numbers, and missing values, we can use the `df_status` function.

```{r}
kable(df_status(df_train_f, FALSE)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

* `q_zeros`: quantity of zeros (`p_zeros`: in percent)
* `q_inf`: quantity of infinite values (`p_inf`: in percent)
* `q_na`: quantity of NA (`p_na`: in percent)
* `type`: factor, ordered-factor, numeric, integer or character
* `unique`: quantity of unique values

### Validation

Let's use `glimpse` function to display a vertical preview of the validation dataset. So we can easily preview data type and sample data.

```{r}
glimpse(df_validate_f)
```

If we want to get some metrics about data types, zeros, infinite numbers, and missing values, we can use the `df_status` function.

```{r}
kable(df_status(df_train_f, FALSE)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

* `q_zeros`: quantity of zeros (`p_zeros`: in percent)
* `q_inf`: quantity of infinite values (`p_inf`: in percent)
* `q_na`: quantity of NA (`p_na`: in percent)
* `type`: factor, ordered-factor, numeric, integer or character
* `unique`: quantity of unique values

### Observations

> Are all the variables in the correct data type?

None. It seems that this has already been dealt with in preprocessing.

> Any variables with lots of zeros?

Yes. Variables with lots of zeros may not be useful for modeling and, in some cases, they may dramatically bias the model. For example, the `content_crashes` is 100% equal to zero.

> Any variables with lots of NAs?

None. Good news.

> Any high cardinality variable?

Factor/categorical variables with a high number of different values (~30) tend to do overfitting if the categories have low cardinality.

## Step 2 : Beta vs Release

### Training

```{r, fig.width = 7.5}
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

f <- freq(df_train_f$label)
```

```{r}
summary(df_release)
```

```{r}
summary(df_beta)
```

### Validation

```{r, fig.width = 7.5}
df_v_release <- df_validate_f[which(df_validate_f$label == 'release'), ]
df_v_beta <- df_validate_f[which(df_validate_f$label == 'beta'), ]

f <- freq(df_validate_f$label)
```

```{r}
summary(df_v_release)
```

```{r}
summary(df_v_beta)
```

## Step 3 - Analyzing Discrete Variables

### Training

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_beta, ggtheme = theme_minimal(base_size = 15))
```

### Validation

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_v_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_v_beta, ggtheme = theme_minimal(base_size = 15))
```

## Step 4 - Analyzing Continuos Variables

### Training

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15))
```

### Validation

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15))
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15))
```

### Observations

* My first impression is that **user engagement** metrics might be a good path to follow

### Ploting Density Curves

```{r include=FALSE}
require(cowplot)
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=uri_count, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 1000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="URI Count", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=uri_count, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 1000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="URI Count", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=active_hours, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Active Hours", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=active_hours, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Active Hours", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=num_pages, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 50000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Num Pages", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=num_pages, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 50000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Num Pages", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=session_length, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 75) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Session Length", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=session_length, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 75) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Session Length", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

# User Engagement Continuous Metrics

This section will focus only on user engagement continuous metrics. So, we are going to analyze the following metrics:

* `num_active_days`
* `active_hours`
* `active_hours_max`
* `uri_count`
* `uri_count_max`
* `session_length`
* `session_length_max`
* `search_count`
* `search_count_max`
* `num_bookmarks`
* `num_pages`
* `num_pages_max`
* `num_addons`
* `daily_unique_domains`
* `daily_unique_domains_max`
* `daily_max_tabs`
* `daily_max_tabs_max`
* `daily_tabs_opened`
* `daily_tabs_opened_max`
* `daily_num_sessions_started`
* `daily_num_sessions_started_max`
* `startup_ms`
* `install_year`
* `profile_age`
* `timezone_offset`
* `memory_mb`
* `cpu_cores`
* `cpu_speed_mhz`
* `cpu_l2_cache_kb`

```{r include=FALSE}
user_eng <- c("num_active_days","active_hours","active_hours_max","uri_count","uri_count_max","session_length","session_length_max","search_count","search_count_max","num_bookmarks","num_pages","num_pages_max","num_addons","daily_unique_domains","daily_unique_domains_max","daily_max_tabs","daily_max_tabs_max","daily_tabs_opened","daily_tabs_opened_max","daily_num_sessions_started","daily_num_sessions_started_max","startup_ms","install_year","profile_age","timezone_offset","memory_mb","cpu_cores","cpu_speed_mhz","cpu_l2_cache_kb")
```

```{r include=FALSE}
df_beta_ue <- df_beta %>% select(user_eng)
df_release_ue <- df_release %>% select(user_eng)
df_train_ue <- df_train_f %>% select(c(user_eng, "label"))

df_beta_v_ue <- df_v_beta %>% select(user_eng)
df_release_v_ue <- df_v_release %>% select(user_eng)
df_validate_ue <- df_validate_f %>% select(c(user_eng, "label"))
```

```{r include=FALSE}
text_tbl <- data.frame(
beta_num_active_days = c(summary(df_beta_ue$num_active_days)), release_num_active_days = c(summary(df_release_ue$num_active_days)), 
beta_active_hours = c(summary(df_beta_ue$active_hours)), release_active_hours = c(summary(df_release_ue$active_hours)), 
beta_active_hours_max = c(summary(df_beta_ue$active_hours_max)), release_active_hours_max = c(summary(df_release_ue$active_hours_max)), 
beta_uri_count = c(summary(df_beta_ue$uri_count)), release_uri_count = c(summary(df_release_ue$uri_count)), 
beta_uri_count_max = c(summary(df_beta_ue$uri_count_max)), release_uri_count_max = c(summary(df_release_ue$uri_count_max)), 
beta_session_length = c(summary(df_beta_ue$session_length)), release_session_length = c(summary(df_release_ue$session_length)), 
beta_session_length_max = c(summary(df_beta_ue$session_length_max)), release_session_length_max = c(summary(df_release_ue$session_length_max)), 
beta_search_count = c(summary(df_beta_ue$search_count)), release_search_count = c(summary(df_release_ue$search_count)), 
beta_search_count_max = c(summary(df_beta_ue$search_count_max)), release_search_count_max = c(summary(df_release_ue$search_count_max)), 
beta_num_bookmarks = c(summary(df_beta_ue$num_bookmarks)), release_num_bookmarks = c(summary(df_release_ue$num_bookmarks)), 
beta_num_pages = c(summary(df_beta_ue$num_pages)), release_num_pages = c(summary(df_release_ue$num_pages)), 
beta_num_pages_max = c(summary(df_beta_ue$num_pages_max)), release_num_pages_max = c(summary(df_release_ue$num_pages_max)), 
beta_daily_unique_domains = c(summary(df_beta_ue$daily_unique_domains)), release_daily_unique_domains = c(summary(df_release_ue$daily_unique_domains)), 
beta_daily_max_tabs = c(summary(df_beta_ue$daily_max_tabs)), release_daily_max_tabs = c(summary(df_release_ue$daily_max_tabs)), 
beta_daily_tabs_opened = c(summary(df_beta_ue$daily_tabs_opened)), release_daily_tabs_opened = c(summary(df_release_ue$daily_tabs_opened)), 
beta_daily_num_sessions_started = c(summary(df_beta_ue$daily_num_sessions_started)), release_daily_num_sessions_started = c(summary(df_release_ue$daily_num_sessions_started)), 
beta_daily_unique_domains_max = c(summary(df_beta_ue$daily_unique_domains_max)), release_daily_unique_domains_max = c(summary(df_release_ue$daily_unique_domains_max)), 
beta_daily_max_tabs_max = c(summary(df_beta_ue$daily_max_tabs_max)), release_daily_max_tabs_max = c(summary(df_release_ue$daily_max_tabs_max)), 
beta_daily_tabs_opened_max = c(summary(df_beta_ue$daily_tabs_opened_max)), release_daily_tabs_opened_max = c(summary(df_release_ue$daily_tabs_opened_max)), 
beta_daily_num_sessions_started_max = c(summary(df_beta_ue$daily_num_sessions_started_max)), release_daily_num_sessions_started_max = c(summary(df_release_ue$daily_num_sessions_started_max)), 
beta_startup_ms = c(summary(df_beta_ue$startup_ms)), release_startup_ms = c(summary(df_release_ue$startup_ms)), 
beta_install_year = c(summary(df_beta_ue$install_year)), release_install_year = c(summary(df_release_ue$install_year)), 
beta_profile_age = c(summary(df_beta_ue$profile_age)), release_profile_age = c(summary(df_release_ue$profile_age)), 
beta_timezone_offset = c(summary(df_beta_ue$timezone_offset)), release_timezone_offset = c(summary(df_release_ue$timezone_offset)), 
beta_memory_mb = c(summary(df_beta_ue$memory_mb)), release_memory_mb = c(summary(df_release_ue$memory_mb)), 
beta_cpu_cores = c(summary(df_beta_ue$cpu_cores)), release_cpu_cores = c(summary(df_release_ue$cpu_cores)), 
beta_cpu_speed_mhz = c(summary(df_beta_ue$cpu_speed_mhz)), release_cpu_speed_mhz = c(summary(df_release_ue$cpu_speed_mhz)), 
beta_cpu_l2_cache_kb = c(summary(df_beta_ue$cpu_l2_cache_kb)), release_cpu_l2_cache_kb = c(summary(df_release_ue$cpu_l2_cache_kb))
)
```

```{r include=FALSE}
text_tbl_v <- data.frame(
beta_num_active_days = c(summary(df_beta_v_ue$num_active_days)), release_num_active_days = c(summary(df_release_v_ue$num_active_days)), 
beta_active_hours = c(summary(df_beta_v_ue$active_hours)), release_active_hours = c(summary(df_release_v_ue$active_hours)), 
beta_active_hours_max = c(summary(df_beta_v_ue$active_hours_max)), release_active_hours_max = c(summary(df_release_v_ue$active_hours_max)), 
beta_uri_count = c(summary(df_beta_v_ue$uri_count)), release_uri_count = c(summary(df_release_v_ue$uri_count)), 
beta_uri_count_max = c(summary(df_beta_v_ue$uri_count_max)), release_uri_count_max = c(summary(df_release_v_ue$uri_count_max)), 
beta_session_length = c(summary(df_beta_v_ue$session_length)), release_session_length = c(summary(df_release_v_ue$session_length)), 
beta_session_length_max = c(summary(df_beta_v_ue$session_length_max)), release_session_length_max = c(summary(df_release_v_ue$session_length_max)), 
beta_search_count = c(summary(df_beta_v_ue$search_count)), release_search_count = c(summary(df_release_v_ue$search_count)), 
beta_search_count_max = c(summary(df_beta_v_ue$search_count_max)), release_search_count_max = c(summary(df_release_v_ue$search_count_max)), 
beta_num_bookmarks = c(summary(df_beta_v_ue$num_bookmarks)), release_num_bookmarks = c(summary(df_release_v_ue$num_bookmarks)), 
beta_num_pages = c(summary(df_beta_v_ue$num_pages)), release_num_pages = c(summary(df_release_v_ue$num_pages)), 
beta_num_pages_max = c(summary(df_beta_v_ue$num_pages_max)), release_num_pages_max = c(summary(df_release_v_ue$num_pages_max)), 
beta_daily_unique_domains = c(summary(df_beta_v_ue$daily_unique_domains)), release_daily_unique_domains = c(summary(df_release_v_ue$daily_unique_domains)), 
beta_daily_max_tabs = c(summary(df_beta_v_ue$daily_max_tabs)), release_daily_max_tabs = c(summary(df_release_v_ue$daily_max_tabs)), 
beta_daily_tabs_opened = c(summary(df_beta_v_ue$daily_tabs_opened)), release_daily_tabs_opened = c(summary(df_release_v_ue$daily_tabs_opened)), 
beta_daily_num_sessions_started = c(summary(df_beta_v_ue$daily_num_sessions_started)), release_daily_num_sessions_started = c(summary(df_release_v_ue$daily_num_sessions_started)), 
beta_daily_unique_domains_max = c(summary(df_beta_v_ue$daily_unique_domains_max)), release_daily_unique_domains_max = c(summary(df_release_v_ue$daily_unique_domains_max)), 
beta_daily_max_tabs_max = c(summary(df_beta_v_ue$daily_max_tabs_max)), release_daily_max_tabs_max = c(summary(df_release_v_ue$daily_max_tabs_max)), 
beta_daily_tabs_opened_max = c(summary(df_beta_v_ue$daily_tabs_opened_max)), release_daily_tabs_opened_max = c(summary(df_release_v_ue$daily_tabs_opened_max)), 
beta_daily_num_sessions_started_max = c(summary(df_beta_v_ue$daily_num_sessions_started_max)), release_daily_num_sessions_started_max = c(summary(df_release_v_ue$daily_num_sessions_started_max)), 
beta_startup_ms = c(summary(df_beta_v_ue$startup_ms)), release_startup_ms = c(summary(df_release_v_ue$startup_ms)), 
beta_install_year = c(summary(df_beta_v_ue$install_year)), release_install_year = c(summary(df_release_v_ue$install_year)), 
beta_profile_age = c(summary(df_beta_v_ue$profile_age)), release_profile_age = c(summary(df_release_v_ue$profile_age)), 
beta_timezone_offset = c(summary(df_beta_v_ue$timezone_offset)), release_timezone_offset = c(summary(df_release_v_ue$timezone_offset)), 
beta_memory_mb = c(summary(df_beta_v_ue$memory_mb)), release_memory_mb = c(summary(df_release_v_ue$memory_mb)), 
beta_cpu_cores = c(summary(df_beta_v_ue$cpu_cores)), release_cpu_cores = c(summary(df_release_v_ue$cpu_cores)), 
beta_cpu_speed_mhz = c(summary(df_beta_v_ue$cpu_speed_mhz)), release_cpu_speed_mhz = c(summary(df_release_v_ue$cpu_speed_mhz)), 
beta_cpu_l2_cache_kb = c(summary(df_beta_v_ue$cpu_l2_cache_kb)), release_cpu_l2_cache_kb = c(summary(df_release_v_ue$cpu_l2_cache_kb))
)
```

## Training

```{r}
kable(text_tbl) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Validation

```{r}
kable(text_tbl_v) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Comparing Two Continuous Distributions

### Training

The QQ plot can be used to compare two continuous distributions.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng) {
  x <- df_beta_ue[,i]
  y <- df_release_ue[,i]

  rg <- range(x, y, na.rm=T)

  test <- ks.test(x, y)$statistic
  pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), paste(pvalue, "\n", test), adj=c(0,1))
  abline(0,1, col="red")  
}
```

```{r message=FALSE, warning=FALSE, out.width=c('50%', '50%')}
for (i in user_eng) {
  x <- df_beta_ue[,i]
  y <- df_release_ue[,i]

  print(i)
  # print(ad.test(x))
  # print(ad.test(y))
  print(ks.test(x, y))
}
```

### Validation

The QQ plot can be used to compare two continuous distributions.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng) {
  x <- df_beta_v_ue[,i]
  y <- df_release_v_ue[,i]

  rg <- range(x, y, na.rm=T)

  test <- ks.test(x, y)$statistic
  pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), paste(pvalue, "\n", test), adj=c(0,1))
  abline(0,1, col="red")  
}
```

```{r message=FALSE, warning=FALSE, out.width=c('50%', '50%')}
for (i in user_eng) {
  x <- df_beta_v_ue[,i]
  y <- df_release_v_ue[,i]

  print(i)
  # print(ad.test(x))
  # print(ad.test(y))
  print(ks.test(x, y))
}
```

### Observations

* Analyzing only the plots, in general, the distributions are very similar to each other
* Mainly the metrics related to **active hours** and **number of active days**, **search count**, **number of pages**, **daily unique domains** and **daily number of sessions started**
* Comparing only the training and validation plots, we can notice that the variables that presented a more different behavior were: `num_active_days`, `uri_count_max` and `num_addons`
* When I performed the KS (kolmogorov-Smirnov) test, the null hypothesis that both samples come from the same distribution was rejected for all the comparisons. *Am I doing something wrong?*

# User Engagement Discrete Metrics

This section will focus only on user engagement continuous metrics. So, we are going to analyze the following metrics:

* `default_search_engine`
* `is_default_browser`
* `profile_age_cat`
* `distro_id_norm`
* `memory_cat`
* `cpu_speed_cat`
* `cpu_cores_cat`
* `cpu_l2_cache_kb_cat`
* `cpu_vendor`
* `os_version`
* `is_wow64`
* `fxa_configured`
* `sync_configured`
* `locale`
* `country`
* `timezone_cat`
* `label`
* `normalized_channel`
* `is_release`

```{r include=FALSE}
user_eng_dis <- c("default_search_engine","is_default_browser","profile_age_cat","distro_id_norm","memory_cat","cpu_speed_cat","cpu_cores_cat","cpu_l2_cache_kb_cat","cpu_vendor","os_version","is_wow64","fxa_configured","sync_configured","locale","country","timezone_cat","label","normalized_channel","is_release")
```

```{r include=FALSE}
df_beta_ue_dis <- df_beta %>% select(user_eng_dis)
df_release_ue_dis <- df_release %>% select(user_eng_dis)
df_train_ue_dis <- df_train_f %>% select(c(user_eng_dis, "label"))

df_beta_v_ue_dis <- df_v_beta %>% select(user_eng_dis)
df_release_v_ue_dis <- df_v_release %>% select(user_eng_dis)
df_validate_ue_dis <- df_validate_f %>% select(c(user_eng_dis, "label"))
```

## Comparing Two Discrete Distributions

### Training

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng_dis) {
  x <- df_beta_ue_dis[,i]
  y <- df_release_ue_dis[,i]

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

  # Add the text 
  text(release_bar, rel_release+0.025, paste(round(rel_release*100), "%", sep="") ,cex=1) 
}
```

### Validation

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng_dis) {
  x <- df_beta_v_ue_dis[,i]
  y <- df_release_v_ue_dis[,i]

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "navy",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

  # Add the text 
  text(release_bar, rel_release+0.025, paste(round(rel_release*100), "%", sep="") ,cex=1) 
}
```

### Observations

* Analyzing only the plots, in general, the distributions are very similar to each other