### 2020-02-17 poc/matchIt/feature_engineering.Rmd
adding feature engineering and selection

---
title: "Feature Engineering"
author: "Mariana de Oliveira Santos Silva"
date: 'Last Updated: `r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: "flatly"
    toc: true
    toc_float: true
    code_folding: "hide"
    highlight: tango
---

# tl;dr

Often combining variables to generate a new variable, a process known as *feature engineering*, can make a model infinitely better. In this notebook we will try to engineer some features in case some of them turn out to be determinant factors when imputing other variables.

```{r imports, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)      # help you build common complex tables and manipulate table styles
library(DataExplorer)    # automated data exploration
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(ggthemes)
library(tidyr)
library(tidyselect)
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(vtreat)          # to variable preparation for ML
library(semTools)        # useful tools for structural equation modeling
library(onehot)
library(mltools)
library(data.table)
load("~/GitHub/ff-beta-release-matching/poc/EDA/data_milestone2_df_train_validate_20191025.RData")
```

```{r echo=FALSE}
## Examine data briefly
glimpse(df_train_f)
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_train_f, ggtheme = theme_minimal(base_size = 15), title = 'Categorical Variables')
```

# Converting Categorical Features

* Most values of `default_search_engine` are equal to **Google**. That is, for better modeling, we could just create a variable that represents *Google* and set the remainder to *Other*. 
* The same can be done for `distro_id_norm`. We could just create a variable that represents *Mozilla* and set the remainder to *Other*.

```{r}
# Training
df_train_f$google_default <- (df_train_f$default_search_engine == 'Google') * 1
df_train_f$mozilla_distro <- (df_train_f$distro_id_norm == 'Mozilla') * 1

# Validation
df_validate_f$google_default <- (df_validate_f$default_search_engine == 'Google') * 1
df_validate_f$mozilla_distro <- (df_validate_f$distro_id_norm == 'Mozilla') * 1
```

```{r}
plot_bar(df_train_f$google_default, ggtheme = theme_minimal(base_size = 15), title = 'Google default seach engine')
plot_bar(df_train_f$mozilla_distro, ggtheme = theme_minimal(base_size = 15), title = 'Mozilla distro id norm')
```

* In addition, we could map numerical values to the categorical variables.

```{r message=FALSE, warning=FALSE}
require(plyr) # to use mapvalues function

# Training
df_train_f$new_os_version <- mapvalues(df_train_f$os_version,
          from=c('Other', '10.0', '6.1', '6.3', '6.2'), 
          to=c(0,1,2,3,4))
df_train_f$new_os_version <- as.numeric(as.character(df_train_f$new_os_version))

df_train_f$new_cpu_vendor <- mapvalues(df_train_f$cpu_vendor,
          from=c('Other', 'Intel', 'AMD'), 
          to=c(0,1,2))
df_train_f$new_cpu_vendor <- as.numeric(as.character(df_train_f$new_cpu_vendor))

# Validation
df_validate_f$new_os_version <- mapvalues(df_validate_f$os_version,
          from=c('Other', '10.0', '6.1', '6.3', '6.2'), 
          to=c(0,1,2,3,4))
df_validate_f$new_os_version <- as.numeric(as.character(df_validate_f$new_os_version))

df_validate_f$new_cpu_vendor <- mapvalues(df_validate_f$cpu_vendor,
          from=c('Other', 'Intel', 'AMD'), 
          to=c(0,1,2))
df_validate_f$new_cpu_vendor <- as.numeric(as.character(df_validate_f$new_cpu_vendor))
```

```{r}
plot_histogram(df_train_f$new_os_version, ggtheme = theme_minimal(base_size = 15), title = 'OS Version')
plot_histogram(df_train_f$new_cpu_vendor, ggtheme = theme_minimal(base_size = 15), title = 'CPU Vendor')
```

# One Hot Encoding

To convert all the nominal or factor variables to numeric, we need to create a dummy variable for each category of the categorical variables. Only one dummy variable is coded with a one for each set of categories.

## Training
```{r}
df_train_encoder <- one_hot(as.data.table(df_train_f))
```

```{r}
numericVarNames <- which(sapply(df_train_encoder, is.numeric))
df_num <- df_train_encoder %>% select(numericVarNames)
df_cat <- df_train_encoder  %>% select(-numericVarNames)

cat('There are', length(df_num), 'numeric variables, and', length(df_cat), 'factor variables')
```

## Validation
```{r}
df_validate_encoder <- one_hot(as.data.table(df_validate_f))
```

```{r}
numericVarNames <- which(sapply(df_validate_encoder, is.numeric))
df_num <- df_validate_encoder %>% select(numericVarNames)
df_cat <- df_validate_encoder  %>% select(-numericVarNames)

cat('There are', length(df_num), 'numeric variables, and', length(df_cat), 'factor variables')
```

# Saving

```{r}
save.image(file = "feature_engineering.RData")
```