### 2020-04-09 poc/EDA/EDA.Rmd

---
title: "Mozilla Data Science Project - 2019"
author: "Mariana de Oliveira Santos Silva"
date: "December 05, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    df_print: kable
    code_folding: "hide"
    theme: "flatly"
---

# Introduction

The goal of this project is to utilize [statistical matching](https://en.wikipedia.org/wiki/Matching_(statistics)) techniques to search for a subset of the prerelease user population that is most representative of Release.

In this first report, we will perform an exploratory analysis of the data, focusing on investigating the differences between Beta and Release users.

# Loading the data

```{r message=FALSE, warning=FALSE, include=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
library(emdist)
library(ggridges)
library(cowplot)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Loading the training dataset

load("~/GitHub/ff-beta-release-matching/poc/EDA/data_milestone2_df_train_validate_20191025.RData")
```

## Training

* App Version 67

```{r message=FALSE, warning=FALSE}
## View train dataframe

kable(head(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Validation

* App Version 68

```{r message=FALSE, warning=FALSE}
## View train dataframe

kable(head(df_validate_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

# Data Inspection

## Training

To get introduced to our training dataset, let's have a look on the basic information of the dataset.

```{r echo=FALSE}
kable(introduce(df_train_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

## Validation

To get introduced to our validation dataset, let's have a look on the basic information of the dataset.

```{r echo=FALSE}
kable(introduce(df_validate_f)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

------

## Observations

```{r echo=FALSE, fig.height=10, fig.width=10}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

plot_intro(df_train_f, ggtheme = theme_minimal(), title = "Training")
plot_intro(df_validate_f, ggtheme = theme_minimal(), title = "Validation")
```

* In both datasets, most variables are continuous
* No NAs value (were handled in preprocessing)

# Data Structure

## Training

Let's use `glimpse` function to display a vertical preview of the training dataset. So we can easily preview data type and sample data.

```{r}
glimpse(df_train_f)
```

If we want to get some metrics about data types, zeros, infinite numbers, and missing values, we can use the `df_status` function.

```{r}
kable(df_status(df_train_f, FALSE)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

* `q_zeros`: quantity of zeros (`p_zeros`: in percent)
* `q_inf`: quantity of infinite values (`p_inf`: in percent)
* `q_na`: quantity of NA (`p_na`: in percent)
* `type`: factor, ordered-factor, numeric, integer or character
* `unique`: quantity of unique values

## Validation

Let's use `glimpse` function to display a vertical preview of the validation dataset.

```{r}
glimpse(df_validate_f)
```

```{r}
kable(df_status(df_train_f, FALSE)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

------

## Observations

> Are all the variables in the correct data type?

None. It seems that this has already been dealt with in preprocessing.

> Any variables with lots of zeros?

Yes. Variables with lots of zeros may not be useful for modeling and, in some cases, they may dramatically bias the model. For example, the `content_crashes` is 100% equal to zero.

> Any variables with lots of NAs?

None. Good news.

> Any high cardinality variable?

Factor/categorical variables with a high number of different values (~30) tend to do overfitting if the categories have low cardinality.

# Beta vs Release

```{r}
## Training
df_release <- df_train_f[which(df_train_f$label == 'release'), ]
df_beta <- df_train_f[which(df_train_f$label == 'beta'), ]

## Validation
df_v_release <- df_validate_f[which(df_validate_f$label == 'release'), ]
df_v_beta <- df_validate_f[which(df_validate_f$label == 'beta'), ]
```

## Training

* Composition:

```{r}
f <- freq(df_train_f$label)
```

## Validation

* Composition:

```{r}
f <- freq(df_validate_f$label)
```

# Analyzing Discrete Variables

## Training

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

## Validation

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution release dataframe
plot_bar(df_v_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
## Frequency distribution beta dataframe
plot_bar(df_v_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

# Analyzing Continuos Variables

## Training

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

## Validation

```{r, fig.width = 15, fig.height = 15}
## View histogram of release dataset
plot_histogram(df_release, ggtheme = theme_minimal(base_size = 15), title = 'Release')
```

```{r, fig.width = 15, fig.height = 15}
## View histogram of beta dataset
plot_histogram(df_beta, ggtheme = theme_minimal(base_size = 15), title = 'Beta')
```

------

## Observations

* My first impression is that **user engagement** metrics might be a good path to follow

## Ploting Density Curves

```{r message=FALSE, warning=FALSE, include=FALSE}
require(cowplot)
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=uri_count, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 1000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="URI Count", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=uri_count, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 1000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="URI Count", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=active_hours, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Active Hours", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=active_hours, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Active Hours", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=num_pages, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 50000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Num Pages", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=num_pages, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 50000) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Num Pages", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

```{r, fig.width = 20, fig.height = 5, message=FALSE, warning=FALSE}
## Training
t <- ggplot(data=df_train_f, aes(x=session_length, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 75) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Session Length", y = "Density") +
    theme_ipsum()

## Validation
v <- ggplot(data=df_validate_f, aes(x=session_length, group=label, fill=label)) +
    geom_density(adjust=1.5, alpha=0.6) + xlim(0, 75) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    labs(x="Session Length", y = "Density") +
    theme_ipsum()

plot_grid(t, v, ncol=2, labels = c("Train", "Validate")) ## Set up a 2 x 2 plotting space
```

# User Engagement Metrics

## Continuous Variables

This section will focus only on user engagement continuous metrics. So, we are going to analyze the following metrics:

* `num_active_days`
* `active_hours`
* `active_hours_max`
* `uri_count`
* `uri_count_max`
* `session_length`
* `session_length_max`
* `search_count`
* `search_count_max`
* `num_bookmarks`
* `num_pages`
* `num_pages_max`
* `num_addons`
* `daily_unique_domains`
* `daily_unique_domains_max`
* `daily_max_tabs`
* `daily_max_tabs_max`
* `daily_tabs_opened`
* `daily_tabs_opened_max`
* `daily_num_sessions_started`
* `daily_num_sessions_started_max`
* `startup_ms`
* `install_year`
* `profile_age`
* `timezone_offset`
* `memory_mb`
* `cpu_cores`
* `cpu_speed_mhz`
* `cpu_l2_cache_kb`

```{r include=FALSE}
user_eng <- c("num_active_days","active_hours","active_hours_max","uri_count","uri_count_max","session_length","session_length_max","search_count","search_count_max","num_bookmarks","num_pages","num_pages_max","num_addons","daily_unique_domains","daily_unique_domains_max","daily_max_tabs","daily_max_tabs_max","daily_tabs_opened","daily_tabs_opened_max","daily_num_sessions_started","daily_num_sessions_started_max","startup_ms","install_year","profile_age","timezone_offset","memory_mb","cpu_cores","cpu_speed_mhz","cpu_l2_cache_kb")
```

```{r include=FALSE}
df_beta_ue <- df_beta %>% select(user_eng)
df_release_ue <- df_release %>% select(user_eng)
df_train_ue <- df_train_f %>% select(c(user_eng, "label"))

df_beta_v_ue <- df_v_beta %>% select(user_eng)
df_release_v_ue <- df_v_release %>% select(user_eng)
df_validate_ue <- df_validate_f %>% select(c(user_eng, "label"))
```

```{r include=FALSE}
text_tbl <- data.frame(
beta_num_active_days = c(summary(df_beta_ue$num_active_days)), release_num_active_days = c(summary(df_release_ue$num_active_days)), 
beta_active_hours = c(summary(df_beta_ue$active_hours)), release_active_hours = c(summary(df_release_ue$active_hours)), 
beta_active_hours_max = c(summary(df_beta_ue$active_hours_max)), release_active_hours_max = c(summary(df_release_ue$active_hours_max)), 
beta_uri_count = c(summary(df_beta_ue$uri_count)), release_uri_count = c(summary(df_release_ue$uri_count)), 
beta_uri_count_max = c(summary(df_beta_ue$uri_count_max)), release_uri_count_max = c(summary(df_release_ue$uri_count_max)), 
beta_session_length = c(summary(df_beta_ue$session_length)), release_session_length = c(summary(df_release_ue$session_length)), 
beta_session_length_max = c(summary(df_beta_ue$session_length_max)), release_session_length_max = c(summary(df_release_ue$session_length_max)), 
beta_search_count = c(summary(df_beta_ue$search_count)), release_search_count = c(summary(df_release_ue$search_count)), 
beta_search_count_max = c(summary(df_beta_ue$search_count_max)), release_search_count_max = c(summary(df_release_ue$search_count_max)), 
beta_num_bookmarks = c(summary(df_beta_ue$num_bookmarks)), release_num_bookmarks = c(summary(df_release_ue$num_bookmarks)), 
beta_num_pages = c(summary(df_beta_ue$num_pages)), release_num_pages = c(summary(df_release_ue$num_pages)), 
beta_num_pages_max = c(summary(df_beta_ue$num_pages_max)), release_num_pages_max = c(summary(df_release_ue$num_pages_max)), 
beta_daily_unique_domains = c(summary(df_beta_ue$daily_unique_domains)), release_daily_unique_domains = c(summary(df_release_ue$daily_unique_domains)), 
beta_daily_max_tabs = c(summary(df_beta_ue$daily_max_tabs)), release_daily_max_tabs = c(summary(df_release_ue$daily_max_tabs)), 
beta_daily_tabs_opened = c(summary(df_beta_ue$daily_tabs_opened)), release_daily_tabs_opened = c(summary(df_release_ue$daily_tabs_opened)), 
beta_daily_num_sessions_started = c(summary(df_beta_ue$daily_num_sessions_started)), release_daily_num_sessions_started = c(summary(df_release_ue$daily_num_sessions_started)), 
beta_daily_unique_domains_max = c(summary(df_beta_ue$daily_unique_domains_max)), release_daily_unique_domains_max = c(summary(df_release_ue$daily_unique_domains_max)), 
beta_daily_max_tabs_max = c(summary(df_beta_ue$daily_max_tabs_max)), release_daily_max_tabs_max = c(summary(df_release_ue$daily_max_tabs_max)), 
beta_daily_tabs_opened_max = c(summary(df_beta_ue$daily_tabs_opened_max)), release_daily_tabs_opened_max = c(summary(df_release_ue$daily_tabs_opened_max)), 
beta_daily_num_sessions_started_max = c(summary(df_beta_ue$daily_num_sessions_started_max)), release_daily_num_sessions_started_max = c(summary(df_release_ue$daily_num_sessions_started_max)), 
beta_startup_ms = c(summary(df_beta_ue$startup_ms)), release_startup_ms = c(summary(df_release_ue$startup_ms)), 
beta_install_year = c(summary(df_beta_ue$install_year)), release_install_year = c(summary(df_release_ue$install_year)), 
beta_profile_age = c(summary(df_beta_ue$profile_age)), release_profile_age = c(summary(df_release_ue$profile_age)), 
beta_timezone_offset = c(summary(df_beta_ue$timezone_offset)), release_timezone_offset = c(summary(df_release_ue$timezone_offset)), 
beta_memory_mb = c(summary(df_beta_ue$memory_mb)), release_memory_mb = c(summary(df_release_ue$memory_mb)), 
beta_cpu_cores = c(summary(df_beta_ue$cpu_cores)), release_cpu_cores = c(summary(df_release_ue$cpu_cores)), 
beta_cpu_speed_mhz = c(summary(df_beta_ue$cpu_speed_mhz)), release_cpu_speed_mhz = c(summary(df_release_ue$cpu_speed_mhz)), 
beta_cpu_l2_cache_kb = c(summary(df_beta_ue$cpu_l2_cache_kb)), release_cpu_l2_cache_kb = c(summary(df_release_ue$cpu_l2_cache_kb))
)
```

```{r include=FALSE}
text_tbl_v <- data.frame(
beta_num_active_days = c(summary(df_beta_v_ue$num_active_days)), release_num_active_days = c(summary(df_release_v_ue$num_active_days)), 
beta_active_hours = c(summary(df_beta_v_ue$active_hours)), release_active_hours = c(summary(df_release_v_ue$active_hours)), 
beta_active_hours_max = c(summary(df_beta_v_ue$active_hours_max)), release_active_hours_max = c(summary(df_release_v_ue$active_hours_max)), 
beta_uri_count = c(summary(df_beta_v_ue$uri_count)), release_uri_count = c(summary(df_release_v_ue$uri_count)), 
beta_uri_count_max = c(summary(df_beta_v_ue$uri_count_max)), release_uri_count_max = c(summary(df_release_v_ue$uri_count_max)), 
beta_session_length = c(summary(df_beta_v_ue$session_length)), release_session_length = c(summary(df_release_v_ue$session_length)), 
beta_session_length_max = c(summary(df_beta_v_ue$session_length_max)), release_session_length_max = c(summary(df_release_v_ue$session_length_max)), 
beta_search_count = c(summary(df_beta_v_ue$search_count)), release_search_count = c(summary(df_release_v_ue$search_count)), 
beta_search_count_max = c(summary(df_beta_v_ue$search_count_max)), release_search_count_max = c(summary(df_release_v_ue$search_count_max)), 
beta_num_bookmarks = c(summary(df_beta_v_ue$num_bookmarks)), release_num_bookmarks = c(summary(df_release_v_ue$num_bookmarks)), 
beta_num_pages = c(summary(df_beta_v_ue$num_pages)), release_num_pages = c(summary(df_release_v_ue$num_pages)), 
beta_num_pages_max = c(summary(df_beta_v_ue$num_pages_max)), release_num_pages_max = c(summary(df_release_v_ue$num_pages_max)), 
beta_daily_unique_domains = c(summary(df_beta_v_ue$daily_unique_domains)), release_daily_unique_domains = c(summary(df_release_v_ue$daily_unique_domains)), 
beta_daily_max_tabs = c(summary(df_beta_v_ue$daily_max_tabs)), release_daily_max_tabs = c(summary(df_release_v_ue$daily_max_tabs)), 
beta_daily_tabs_opened = c(summary(df_beta_v_ue$daily_tabs_opened)), release_daily_tabs_opened = c(summary(df_release_v_ue$daily_tabs_opened)), 
beta_daily_num_sessions_started = c(summary(df_beta_v_ue$daily_num_sessions_started)), release_daily_num_sessions_started = c(summary(df_release_v_ue$daily_num_sessions_started)), 
beta_daily_unique_domains_max = c(summary(df_beta_v_ue$daily_unique_domains_max)), release_daily_unique_domains_max = c(summary(df_release_v_ue$daily_unique_domains_max)), 
beta_daily_max_tabs_max = c(summary(df_beta_v_ue$daily_max_tabs_max)), release_daily_max_tabs_max = c(summary(df_release_v_ue$daily_max_tabs_max)), 
beta_daily_tabs_opened_max = c(summary(df_beta_v_ue$daily_tabs_opened_max)), release_daily_tabs_opened_max = c(summary(df_release_v_ue$daily_tabs_opened_max)), 
beta_daily_num_sessions_started_max = c(summary(df_beta_v_ue$daily_num_sessions_started_max)), release_daily_num_sessions_started_max = c(summary(df_release_v_ue$daily_num_sessions_started_max)), 
beta_startup_ms = c(summary(df_beta_v_ue$startup_ms)), release_startup_ms = c(summary(df_release_v_ue$startup_ms)), 
beta_install_year = c(summary(df_beta_v_ue$install_year)), release_install_year = c(summary(df_release_v_ue$install_year)), 
beta_profile_age = c(summary(df_beta_v_ue$profile_age)), release_profile_age = c(summary(df_release_v_ue$profile_age)), 
beta_timezone_offset = c(summary(df_beta_v_ue$timezone_offset)), release_timezone_offset = c(summary(df_release_v_ue$timezone_offset)), 
beta_memory_mb = c(summary(df_beta_v_ue$memory_mb)), release_memory_mb = c(summary(df_release_v_ue$memory_mb)), 
beta_cpu_cores = c(summary(df_beta_v_ue$cpu_cores)), release_cpu_cores = c(summary(df_release_v_ue$cpu_cores)), 
beta_cpu_speed_mhz = c(summary(df_beta_v_ue$cpu_speed_mhz)), release_cpu_speed_mhz = c(summary(df_release_v_ue$cpu_speed_mhz)), 
beta_cpu_l2_cache_kb = c(summary(df_beta_v_ue$cpu_l2_cache_kb)), release_cpu_l2_cache_kb = c(summary(df_release_v_ue$cpu_l2_cache_kb))
)
```

### Training

#### Beta-Release Difference

```{r}
kable(text_tbl) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

### Validation

#### Beta-Release Difference

```{r}
kable(text_tbl_v) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

### QQ-Plots

```{r fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
par(mfrow = c(4, 2))  ## Set up a 2 x 3 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng) {

  # Training
  x_t <- df_beta_ue[,i]
  y_t <- df_release_ue[,i]

  rg_t <- range(x_t, y_t, na.rm=T)

  test_t <- ks.test(x_t, y_t)$statistic
  test_t <- paste("KS Test = ", round(test_t, 3))

  # Validation

  x_v <- df_beta_v_ue[,i]
  y_v <- df_release_v_ue[,i]

  rg_v <- range(x_v, y_v, na.rm=T)

  test_v <- ks.test(x_v, y_v)$statistic
  test_v <- paste("KS Test = ", round(test_v, 3))

  ########

  title_t <- paste('V67', i, sep='\n')
  qqplot(x_t, y_t, main=title_t, xlim=rg_t, ylim=rg_t, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x_t), (if(max(x_t) > max(y_t)) max(x_t) else max(y_t)), test_t, adj=c(0,1))
  abline(0,1, col="#fe346e", lty=2)

  title_v <- paste('V68', i, sep='\n')
  qqplot(x_v, y_v, main=title_v, xlim=rg_v, ylim=rg_v, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x_v), (if(max(x_v) > max(y_v)) max(x_v) else max(y_v)), test_v, adj=c(0,1))
  abline(0,1, col="#fe346e", lty=2)  
}
```

### Observations

```{r message=FALSE, warning=FALSE, include=FALSE}
tbl_ks <- data.frame(
training = c(0.071,0.045,0.044,0.047,0.054,0.099,0.073,0.014,0.019,0.026,0.055,0.054,0.601,0.029,0.033,0.094,0.080,0.035,0.032,0.119,0.116,0.129,0.018,0.042,0.208,0.075,0.064,0.046,0.025), 
validation = c(0.166,0.058,0.077,0.071,0.095,0.076,0.040,0.045,0.059,0.042,0.082,0.082,0.292,0.040,0.054,0.075,0.051,0.036,0.053,0.098,0.117,0.130,0.017,0.035,0.235,0.109,0.087,0.062,0.035)
)

row.names(tbl_ks) <- c('num_active_days','active_hours','active_hours_max','uri_count','uri_count_max','session_length','session_length_max','search_count','search_count_max','num_bookmarks','num_pages','num_pages_max','num_addons','daily_unique_domains','daily_unique_domains_max','daily_max_tabs','daily_max_tabs_max','daily_tabs_opened','daily_tabs_opened_max','daily_num_sessions_started','daily_num_sessions_started_max','startup_ms','install_year','profile_age','timezone_offset','memory_mb','cpu_cores','cpu_speed_mhz','cpu_l2_cache_kb')
```

```{r}
kable(tbl_ks) %>%
add_header_above(c("KS Distance" = 3)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
row_spec(c(2,18,21:26), bold = T, color = "white", background = "#c3f584")
```

* Analyzing at the plots of each dataset separately, in general, Beta-Release distributions are very similar to each other
* For instance, metrics related to **active hours** and **number of active days**, **search count**, **number of pages**, **daily unique domains** and **daily number of sessions started**
* Comparing only the training and validation plots, we can observe large differences between the variables. Only a few showed similar distances (marked in green)

### Violin Plots

The following violin plots depicts distributions for the beta and release subsets, for both versions v67 and v68. Violin plot is a powerful data visualization technique since it allows to compare both the **ranking** of several groups and their **distribution**.

**NOTE:** Guiding lines have been added for the following:

* <span style="color:red">red</span> dashed: Release mean
* <span style="color:blue">blue</span> dashed: Beta mean
* <span style="color:red">red</span> solid: Release median
* <span style="color:blue">blue</span> solid: Beta median

```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}

## Violin plots
for (i in user_eng) {
  df_validate_means <- df_validate_ue %>%
  group_by(label) %>%
  summarise(value = mean(eval(as.name(i))))

  df_validate_medians <- df_validate_ue %>%
  group_by(label) %>%
  summarise(value = median(eval(as.name(i))))

  df_train_means <- df_train_ue %>%
  group_by(label) %>%
  summarise(value = mean(eval(as.name(i))))

  df_train_medians <- df_train_ue %>%
  group_by(label) %>%
  summarise(value = median(eval(as.name(i))))

  ###########

  violin_train <- ggplot(df_train_ue, aes(x=label, y=eval(as.name(i)), fill=label)) + 
  geom_violin(trim=FALSE) +
  labs(title=i,x="Channel", y = "Measure") + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  geom_hline(
    data = df_train_means,
    aes(yintercept = value, colour = label),
    linetype = "dashed",
    size = 1
  ) +
  geom_hline(
    data = df_train_medians,
    aes(yintercept = value, colour = label),
    linetype = "solid",
    size = 1
  ) + scale_colour_manual(values=c("blue", "red")) +
  geom_boxplot(width=0.1, fill="white", alpha=0.5) + 
  theme_minimal() + theme(legend.position="none") + coord_flip()

  violin_valid <- ggplot(df_validate_ue, aes(x=label, y=eval(as.name(i)), fill=label)) + 
  geom_violin(trim=FALSE) +
  labs(title=i,x="Channel", y = "Measure") + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  geom_hline(
    data = df_validate_means,
    aes(yintercept = value, colour = label),
    linetype = "dashed",
    size = 1
  ) +
  geom_hline(
    data = df_validate_medians,
    aes(yintercept = value, colour = label),
    linetype = "solid",
    size = 1
  ) + scale_colour_manual(values=c("blue", "red")) +
  geom_boxplot(width=0.1, fill="white", alpha=0.5) + 
  theme_minimal() + theme(legend.position="none") + coord_flip()

  print(plot_grid(violin_train, violin_valid, ncol = 1,labels = c('V67','V68')))
}
```

### Observations

The violin plots above show the relationship of channel type to user engagement metrics. Overall, the results were quite similar comparing the two versions (v67 and v68). Only the following metrics yielded different results:

* `active_hours_max`
* `daily_unique_domains_max`

Now, let's take a closer look at the similar comparative results:

> Which variables presented **similar/equal** means and medians for beta and release users?

* `uri_count`
* `uri_count_max`
* `session_length_max`
* `search_count`
* `search_count_max`
* `daily_unique_domains`
* `daily_tabs_opened_max`
* `install_year`
* `profile_age`
* `memory_mb`
* `cpu_speed_mhz`

> Which variables presented **higher** release users means and/or medians than betas?

* `num_active_days`
* `active_hours`
* `num_pages`
* `num_pages_max`
* `daily_num_sessions_started`
* `daily_num_sessions_started_max`
* `cpu_cores`

> Which variables presented **lower** release users means and/or medians than betas?

* `session_length`
* `num_addons`
* `daily_tabs_opened`
* `timezone_offset`
* `cpu_l2_cache_kb`

Four variables became difficult to visualize through the violin plot. So we will plot ridgeline charts.

```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
vlist <- c('num_bookmarks','daily_max_tabs','daily_max_tabs_max','startup_ms')

## Ridgeline  plots
for (i in vlist) {
  ridge_train <- ggplot(df_train_ue, aes(x = eval(as.name(i)), y = label, fill = label)) +
  geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
  labs(title=i,x="Channel", y = "Measure") + 
  theme_ridges() + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  theme(legend.position = "none")

  ridge_valid <- ggplot(df_validate_ue, aes(x = eval(as.name(i)), y = label, fill = label)) +
  geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
  labs(title=i,x="Channel", y = "Measure") + 
  theme_ridges() + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  theme(legend.position = "none")

  print(plot_grid(ridge_train, ridge_valid, ncol = 1,labels = c('V67','V68')))
}
```

## Discrete Variables

This section will focus only on user engagement discrete metrics. So, we are going to analyze the following metrics:

* `default_search_engine`
* `is_default_browser`
* `profile_age_cat`
* `distro_id_norm`
* `memory_cat`
* `cpu_speed_cat`
* `cpu_cores_cat`
* `cpu_l2_cache_kb_cat`
* `cpu_vendor`
* `os_version`
* `is_wow64`
* `fxa_configured`
* `sync_configured`
* `locale`
* `country`
* `timezone_cat`
* `label`
* `normalized_channel`
* `is_release`

```{r include=FALSE}
user_eng_dis <- c("default_search_engine","is_default_browser","profile_age_cat","distro_id_norm","memory_cat","cpu_speed_cat","cpu_cores_cat","cpu_l2_cache_kb_cat","cpu_vendor","os_version","is_wow64","fxa_configured","sync_configured","locale","country","timezone_cat","label","normalized_channel","is_release")
```

```{r include=FALSE}
df_beta_ue_dis <- df_beta %>% select(user_eng_dis)
df_release_ue_dis <- df_release %>% select(user_eng_dis)
df_train_ue_dis <- df_train_f %>% select(c(user_eng_dis, "label"))

df_beta_v_ue_dis <- df_v_beta %>% select(user_eng_dis)
df_release_v_ue_dis <- df_v_release %>% select(user_eng_dis)
df_validate_ue_dis <- df_validate_f %>% select(c(user_eng_dis, "label"))
```

### Training

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng_dis) {
  x <- df_beta_ue_dis[,i]
  y <- df_release_ue_dis[,i]

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#111d5e",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#b21f66",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

  # Add the text 
  text(release_bar, rel_release+0.025, paste(round(rel_release*100), "%", sep="") ,cex=1) 
}
```

### Validation

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))  ## Set up a 2 x 2 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng_dis) {
  x <- df_beta_v_ue_dis[,i]
  y <- df_release_v_ue_dis[,i]

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#111d5e",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#b21f66",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

  # Add the text 
  text(release_bar, rel_release+0.025, paste(round(rel_release*100), "%", sep="") ,cex=1) 
}
```

------

### Observations

* In general, the distributions are very similar to each other for both versions