### 2020-04-09 poc/matchIt/model_selection.Rmd

---
title: "Beta to Release Matching: Model Selection"
author: "Mariana de Oliveira Santos Silva"
date: 'Last Updated: `r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: "flatly"
    toc: true
    toc_float: true
    code_folding: "hide"
    highlight: tango
---

```{r include=FALSE}
generate_formula <- function(tr_cov, label){
  # Focus on linear and no-interactions
  return(as.formula(paste(label, '~', paste(tr_cov, collapse="+"))))
}
```

# Introduction

The main objective of this notebook is to perform several different matching methods in experiment 3 (which was the experiment that presented the best results in the previous work) to select the best model.

# Loading the data

```{r message=FALSE, warning=FALSE, include=FALSE}
## Loading the needed libraries

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(DataExplorer)    # automated data exploration
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(dplyr)           # for data frame manipulation
library(MatchIt)
library(optmatch)
library(cem)
library(RItools)
library(magrittr)
library(tidyr)
library(tidyselect)
library(grt)
library(tableone)
library(cobalt)

label <- 'is_release'
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Loading the training dataset
load('~/ff-beta-release-matching/poc/matchIt/modeling.RData')
```

# Model Selection

In short, the prior results showed that the model using *Nearest Neighbor matching* with stratified sampling outperformed the other models. In particular, the `experiment 3` (regarding to the results of feature selection with stratified sampling, considering the top 5 features) presented the best results for most of the tested models. However, the results could be further improved. So, before move on to the last phase (the model validation), we can try using other matching strategies. Now that we have selected the best expirement.

## Match using Coarsened Exact Matching (CEM): Small Dataset

```{r message=FALSE, warning=FALSE}
cem <- matchit(formula = generate_formula(exp_3, label), df_1x_sm, 'cem')
res <- match.data(cem)
cem_models_1x <- cem
cem_results_1x <- res
print(summary(cem))
```

## Match using Coarsened Exact Matching (CEM): Full Dataset

```{r message=FALSE, warning=FALSE}
cem_full <- matchit(formula = generate_formula(exp_3, label), df_1x, 'cem')
res_full <- match.data(cem_full)
cem_full_models_1x <- cem_full
cem_full_results_1x <- res_full
print(summary(cem_full))
```

## Match using Nearest Neighbor matching: Small Dataset

```{r message=FALSE, warning=FALSE}
nn <- matchit(formula = generate_formula(exp_3, label), df_1x_sm, 'nearest', replace = TRUE)
res <- match.data(nn)
nn_models_1x <- nn
nn_results_1x <- res
print(summary(nn))
```

## Match using Nearest Neighbor matching: Full Dataset

```{r message=FALSE, warning=FALSE}
nn_full <- matchit(formula = generate_formula(exp_3, label), df_1x, 'nearest', replace = TRUE)
res <- match.data(nn_full)
nn_full_models_1x <- nn_full
nn_full_results_1x <- res
print(summary(nn_full))
```

## Match using Nearest Neighbor matching: Small Dataset - Mahalanobis

```{r message=FALSE, warning=FALSE}
nn.ma <- matchit(formula = generate_formula(exp_3, label), df_1x_sm, 'nearest', distance='mahalanobis')
res <- match.data(nn.ma)
nn.ma_models_1x <- nn.ma
nn.ma_results_1x <- res
print(summary(nn.ma))
```

## Match using Nearest Neighbor matching: Full Dataset - Mahalanobis

```{r message=FALSE, warning=FALSE}
nn.ma_full <- matchit(formula = generate_formula(exp_3, label), df_1x, 'nearest', distance='mahalanobis')
res <- match.data(nn.ma_full)
nn.ma_full_models_1x <- nn.ma_full
nn.ma_full_results_1x <- res
print(summary(nn.ma_full))
```

## Match using Subclassification matching: Small Dataset

```{r message=FALSE, warning=FALSE}
sub <- matchit(formula = generate_formula(exp_3, label), df_1x_sm, 'subclass')
res <- match.data(sub)
sub_models_1x <- sub
sub_results_1x <- res
print(summary(sub))
```

## Match using Subclassification matching: Full Dataset

```{r message=FALSE, warning=FALSE}
sub_full <- matchit(formula = generate_formula(exp_3, label), df_1x, 'subclass')
res <- match.data(sub_full)
sub_full_models_1x <- sub_full
sub_full_results_1x <- res
print(summary(sub_full))
```

---

## Graphical Comparison

```{r}
small <- list(cem_models_1x = cem_models_1x, nn_models_1x = nn_models_1x, nn.ma_models_1x = nn.ma_models_1x, sub_models_1x = sub_models_1x)

full <- list(cem_full_models_1x = cem_full_models_1x, nn_full_models_1x = nn_full_models_1x, nn.ma_full_models_1x = nn.ma_full_models_1x, sub_full_models_1x = sub_full_models_1x)
```

* Note that we want to see if the Matched Treated (beta) and Matched Control (release) distributions are roughly similar.

### Small Dataset

```{r fig.height=7, fig.width=14}
for (model in names(small)){
  print(paste(model))
  if (model != "nn.ma_models_1x"){
    plot(small[[model]], type = "hist", col = "#111d5e")
    plot(small[[model]], type = "jitter", col = "#111d5e", interactive = F)
  } else {
    print(paste("Not appropriate for pure Mahalanobis matching. No plots generated."))
  }
}
```

### Full Dataset

```{r fig.height=7, fig.width=14}
for (model in names(full)){
  print(paste(model))
  if (model != "nn.ma_full_models_1x"){
    plot(full[[model]], type = "hist", col = "#111d5e")
    plot(full[[model]], type = "jitter", col = "#111d5e", interactive = F)
  } else {
    print(paste("Not appropriate for pure Mahalanobis matching. No plots generated."))
  }
}
```

---

### Observations

* Although it was not possible to graphically compare the results using the Nearest Neighbor matching with Mahalanobis distance, from the summary, we can see that those models outperformed the others. Analyzing each of the covariates, we have:

#### Small Dataset

```{r fig.height=7, fig.width=14}
for (v in exp_3){
  p <- bal.plot(nn.ma_models_1x, var.name = v, colors = c("#111d5e", "#611e62"))
  print(p)
}
```

#### Full Dataset

```{r fig.height=7, fig.width=14}
for (v in exp_3){
  p <- bal.plot(nn.ma_full_models_1x, var.name = v, colors = c("#111d5e", "#611e62"))
  print(p)
}
```

* The figures show that the model containing the entire dataset was better than sampling. 
* In addition, the main problem is displayed mainly in pageload metrics.